---
title: "Regression Analysis"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
```{r initialize}
source("scripts/0_start.R")
# this already includes lme4, broom.mixed, etc.


regression_packages <- c("VIF", "nlme", "MASS", "spdep")
install_missing_packages(regression_packages)
library(VIF) # for selecting models for a linear model. Recommended by Liang.

# gls model - spatial autocorrelation
# spdep has lots of functions for assessing spatial autocorrelation.


```

# Exponential Decay:


Reminder of the framework for the first paper: characterizing abandonment trajectories & recultivation
- Looking at abandonment length (mean age), persistence, turnover (half-life), and predictors of age and recultivation. To a lesser extent, recultivation rates
- Variation between sites
- Combining the half-life and average length of time abandoned to extrapolate age classes into the future. 
- Possibly including aspects of fragmentation and measures of biomass over time (NDVI) - not likely
- Minor discussion of implications for biodiversity and carbon

Paper 1: characterizing abandonment trajectories (length, persistence, predictors)
1) What proportion of abandoned land remains abandoned long-term, and what proportion is recultivated? 
    a) What are the most common land uses of formerly abandoned lands (does this end up in grassland, forest, neither?)?
2) How long does abandonment typically last for? How much annual (or periodic) turnover occurs?
    a) What is the half-life of abandoned land? How does this vary across site, and overtime at each site?
3) How do these result vary across regions?
4) What are key predictors of length of abandonment, and recultivation? (Slope, elevation, soil, climate variables, surrounding landcover, proximity to woody veg/grassland, population, agricultural marginality)
5) How does landscape fragmentation (of grassland, woody vegetation, and agriculture) change over time as a result of abandonment? To what extent does abandonment result in older and more contiguous grassland and forests?
6) Can we discern regeneration in the trajectory of agriculture -> grassland -> woody vegetation? How does biomass (proxied by NDVI, or otherwise) change overtime? 
5) and 6) can and probably should be moved to paper 2 (biodiversity focus).
- Will include discussion of implications for biodiversity and other environmental factors (e.g. carbon), but based largely on existing data on time to biodiversity recovery.


My main questions: 1) how long does abandoned land stay abandoned for, and 2) how durable is abandoned agricultural land? These in turn translate into a series of more concrete questions: 
1) What is the half-life of abandoned agricultural land? 
2) How does the half-life vary over time?
3) How does the average half-life and the rate of change of the half-life differ across sites?

The plots of persistence of abandoned land over time show this


To provide a simple answer to this, I will first calculate the half-life for each of my sites, then look at how the half-life evolves over time at each of the sites. Therefore, I’m trying to model an exponential decay curve, with the proportion of land remaining abandoned (“proportion") as a function of how long the land has been abandoned (“age”). is the 


```{r load-data}
# load data, produced from the following functions
# cc_calc_persistence() # which is called within
# cc_calc_persistence_all() # which is called within
# cc_summarize_abn_dts() # this is the master call within the cluster script, "analyze_all_sites.R"


site_df <- read.csv(file = paste0(p_dat_derived, "site_df.csv"))


# -------------------------- load summarized data.frames --------------------------- #
indx <- 9
site <- site_df$site[indx] # set site:
site_label <- site_df$label[indx] # set label
blip_label <- "_b1"
load(file = paste0(p_output, "abn_dat_products", blip_label, site_label, ".rds"), verbose = TRUE)

# loads:
area_b1_s
persistence_list_b1_s
abn_area_change_b1_s




# -------------------------- rasters --------------------------- #
# prepared input rasters (derived by Chris)
site_input_raster_files <- list.files(paste0(p_dat_derived, "input_rasters"), full.names = TRUE) %>%
  grep(".tif", ., value = TRUE) #%>% grep("age", ., value = TRUE, invert = TRUE)

r <- lapply(seq_along(site_input_raster_files), function(i) {brick(site_input_raster_files[i])})
names(r) <- site_df$site

# rename raster layers:
for (i in 1:11) {
  if (names(r[i]) == "nebraska") {
    names(r[[i]]) <- paste0("y", 1986:2018)
  } else {
    if (names(r[i]) == "wisconsin") {
      names(r[[i]]) <- paste0("y", 1987:2018)
    } else {
      # everything else, just 1987:2017
      names(r[[i]]) <- paste0("y", 1987:2017)
    }}}

# note: Land use class codes:
#       1. Non-vegetated area (e.g. water, urban, barren land)
#       2. Woody vegetation
#       3. Cropland 
#       4. Herbaceous land (e.g. grassland)

# abandonment age maps (produced by Chris)
age_files <- list.files(paste0(p_dat_derived, "age_rasters"), full.names = TRUE) %>%
  grep(".tif", ., value = TRUE) #%>% grep("age", ., value = TRUE, invert = FALSE)

age_r <- lapply(seq_along(age_files), function(i) {brick(age_files[i])})
names(age_r) <- site_df$site
for (i in seq_along(age_r)) {names(age_r[[i]]) <- paste0("y", 1987:2017)} # remember: these are just 1987:2017

# year of first abandonment maps (from He)
yoa_files <- list.files(paste0(p_dat, "Abandonment/year_of_abandonment/"))



# testers:
bs <- brick(paste0(p_dat, "Abandonment/belarus_small.tif"))
bt <- brick(paste0(p_dat_derived, "belarus_subset.tif"))
names(bs) <- paste0("y", 1987:2017)
names(bt) <- paste0("y", 1987:2017)



# -------------------------- data.tables --------------------------- #
b_age <- fread(input = paste0(p_dat_derived, "belarus_age.csv"))
names(b_age)
s_age <- fread(input = paste0(p_dat_derived, "shaanxi_age.csv"))


b_length <- fread(input = paste0(p_dat_derived, "lengths/", "belarus_length_b1.csv"))
s_length <- fread(input = paste0(p_dat_derived, "lengths/", "shaanxi_length_b1.csv"))

b_max_length <- fread(input = paste0(p_dat_derived, "lengths/", "belarus_max_length_b1.csv"))
s_max_length <- fread(input = paste0(p_dat_derived, "lengths/", "shaanxi_max_length_b1.csv"))

# original data
s_dt <- fread(input = paste0(p_dat_derived, "shaanxi.csv"))
names(s_dt) <- gsub(pattern = "andcover", replacement = "y", names(s_dt))

b_dt <- fread(input = paste0(p_dat_derived, "belarus.csv")) # caution - huge file! 8.4 GB at least. 




```

```{r prep-dat}
str(dat)

blip_label <- "_b1"


pairs(dat)

dat <- persistence_list_b1_s$na_first

dat <- left_join(dat, 
                 dat %>% filter(age == 5) %>% select(year_abn, initial_area_abn = area_ha), 
                 by = "year_abn") %>%
  
  # create bins of year_abn - could also use cut(), e.g. year_bins = cut(year_abn, breaks = 5) 
  # dat %>% mutate(nbins = cut(dat$year_abn, breaks = 5, include.lowest = TRUE)) %>% print(n = 30)
  mutate(year_abn_bins = ifelse(year_abn >= 1988 & year_abn <= 1993, "1988-1993",
                           ifelse(year_abn >= 1994 & year_abn <= 1998, "1994-1998",
                                  ifelse(year_abn >= 1999 & year_abn <= 2003, "1999-2003",
                                         ifelse(year_abn >= 2004 & year_abn <= 2008, "2004-2008",
                                                ifelse(year_abn >= 2009 & year_abn <= 2013, "2009-2013", NA)
                                         ))))) %>%
  mutate(year_abn_bins = as_factor(year_abn_bins), # convert both to factors
         #year_abn = as_factor(year_abn),
         time = age - 5,
         time_1 = time + 1,
         cohort = factor(year_abn)) 




dat0 <- select(dat, time_1, age, initial_area_abn, cohort, year_abn) # for predictions



# variables of interest

dat$proportion # response variable: the proportion of originally abandoned land that remains in a given year. 
dat$year # the current year, allowing us to measure area abandoned in the current year, relative to the amount initially abandoned
dat$year_abn # cohort of land abandoned in a particular year, which is then followed year after year to track persistence
# ^ probably should be a factor
dat$count # number of pixels, from which area is determined
dat$area_ha # area abandoned, of a given cohort
dat$age # how long that land in that cohort has been abandoned for
dat$time # age - 5 
dat$bins # bins of abandoned land of similar ages (5-10 years)
dat$year_abn_bins # bins of land abandoned during the same five to six year window
dat$initial_area_abn # the area (in ha) initially abandoned in a given year (cohort)

dat$time # is just the adjusted time abandoned beyond the initial abandonment threshold. So, 0 corresponds to an abandonment age of 5, 1 -> 6, 2 -> 7 etc.
dat$cohort # a factor to be used for cohort
```

```{r dat-all-sites}
dat_l <- lapply(1:11, FUN = function(x) {
  site <- site_df$site[x] # set site:
  site_label <- site_df$label[x] # set label

  load(file = paste0(p_output, "abn_dat_products", blip_label, site_label, ".rds"), verbose = TRUE)
  assign("dat", eval(parse(text = paste0("persistence_list", blip_label, site_label)))[["na_first"]])
  dat <- left_join(dat,
                   dat %>% filter(age == 5) %>% select(year_abn, initial_area_abn = area_ha), 
                   by = "year_abn") %>%
    mutate(year_abn_bins = cut(year_abn, breaks = 5, include.lowest = TRUE)) %>%
    mutate(year_abn_bins = as_factor(year_abn_bins), # convert both to factors
           time = age - 5,
           time_1 = time + 1,
           cohort = factor(year_abn)) 
  })

names(dat_l) <- site_df$site
dat_l <- dat_l %>% bind_rows(.id = "site")
```


```{r run-lm-all-sites}
# log-log model
lm_log2_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0 + log(time_1):cohort,  
               data = filter(dat_l, site == site_df$site[i]))
})

# lon-linear (exp. decay)
lm_log_lin_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0 + time:cohort,  
               data = filter(dat_l, site == site_df$site[i]))
})

# linear
lm_0_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0 + time,  
               data = filter(dat_l, site == site_df$site[i]))
})

names(lm_log2_l) <- site_df$site
names(lm_log_lin_l) <- site_df$site
names(lm_0_l) <- site_df$site



# save plots:
for (i in 1:11) {
  # set mod:
  mod_tmp <- lm(log(proportion) ~ 0 + log(time_1):cohort,  
               data = filter(dat_l, site == site_df$site[i]))
  
  gg_log2_cohort_tmp <- ggplot(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
    geom_point(data = filter(dat_l, site == site_df$site[i])) + scale_color_distiller(palette = "Greens") +
    scale_x_continuous(n.breaks = 30) + 
    geom_hline(yintercept = 0.5, linetype = "dashed") + 
    labs(title = "Decay of abandoned land over time", 
         subtitle = site_df$description[i], 
         caption = "lm formula: log(proportion) ~ 0 + log(time_1):cohort",
         x = "Time since initial abandonment (years)", 
         y = "Proportion of abandoned land remaining",
         color = "Cohort (Year \nFirst Abandoned)") + 
    
    # fitted values
    geom_line(data = mutate(augment(mod_tmp), 
                          year_abn = as.numeric(levels(cohort))[cohort]),
            aes(x = exp(`log(time_1)`) + 4, y = exp(.fitted))) + 
    
    # average (needs to be transformed 4 to the right, so that plotting works)
    geom_function(fun = function(x) { (x-4) ^ (mean(coef(mod_tmp), na.rm = TRUE))},
                  mapping = aes(linetype = "Mean Decay Rate"),
                  color = "blue", size = 1, inherit.aes = FALSE) + 
    scale_linetype(name = "")
  
  png(filename = paste0(p_output, "plots/log2_cohort", site_df$label[i], ".png"), 
    width = 8, height = 6, units = "in", res = 400)
  print(gg_log2_cohort_tmp)
  dev.off()
  
  
}
```

```{r compare-AIC-all-sites}

lm_glance1 <- lapply(lm_log2_l, FUN = function(x) {
  glance(x)
}) %>% bind_rows(.id = "site") %>% mutate(mod = "log2_cohort") %>% select(site, AIC, mod, everything()) %>% arrange(AIC)

lm_glance2 <- lapply(lm_log_lin_l, FUN = function(x) {
  glance(x)
}) %>% bind_rows(.id = "site") %>% mutate(mod = "log_lin_cohort") %>% select(site, AIC, mod, everything()) %>% arrange(AIC)

lm_glance3 <- lapply(lm_0_l, FUN = function(x) {
  glance(x)
}) %>% bind_rows(.id = "site") %>% mutate(mod = "log_lin_0") %>% select(site, AIC, mod, everything()) %>% arrange(AIC)


lm_g <- bind_rows(lm_glance1, lm_glance2, lm_glance3)

# view results
lm_g %>% arrange(site, AIC) %>% print(n = 33)

# at all sites, the log-log cohort model has the lowest AIC (i.e. best performance)... except Nebraska! Will check that out next.

```

```{r nebraska-log2-v-log-lin}
dat %>% select(-proportion)

gg_nebraska_log2 <- ggplot(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point(data = filter(dat_l, site == "nebraska")) + scale_color_distiller(palette = "Greens") + 
  #scale_x_continuous(n.breaks = 30) + 
  scale_y_continuous(limits = c(0.4, 1)) +
  geom_hline(yintercept = 0.5, linetype = "dashed") + 
  labs(title = "Decay of abandoned land over time", subtitle = site_df$description[7], 
       caption = "lm formula: log(proportion) ~ 0 + log(time_1):cohort",
       x = "Time since initial abandonment (years)", y = "Proportion of abandoned land remaining",
       color = "Cohort (Year \nFirst Abandoned)") + 
  
  # log-log
  geom_line(data = mutate(augment(lm_log2_l$nebraska, 
                                  newdata = filter(dat_l, site == "nebraska") %>% select(-proportion)),
                          proportion = exp(.fitted))) + 
  
  # average (needs to be transformed 4 to the right, so that plotting works)
  geom_function(fun = function(x) { (x-4) ^ (mean(coef(lm_log2_l$nebraska), na.rm = TRUE))},
                mapping = aes(linetype = "Mean Decay Rate"),
                color = "blue", size = 1, #linetype = "dashed", 
                inherit.aes = FALSE) + 
  scale_linetype(name = "")
  

gg_nebraska_log_lin <- ggplot(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point(data = filter(dat_l, site == "nebraska")) + scale_color_distiller(palette = "Greens") + 
  #scale_x_continuous(n.breaks = 30) + 
  scale_y_continuous(limits = c(0.4, 1)) +
  geom_hline(yintercept = 0.5, linetype = "dashed") + 
  labs(title = "Decay of abandoned land over time", subtitle = site_df$description[7], 
       caption = "lm formula: log(proportion) ~ 0 + time:cohort",
       x = "Time since initial abandonment (years)", y = "Proportion of abandoned land remaining",
       color = "Cohort (Year \nFirst Abandoned)") + 
  
  # log-lin
  geom_line(data = mutate(augment(lm_log_lin_l$nebraska, 
                                  newdata = filter(dat_l, site == "nebraska") %>% select(-proportion)),
                          proportion = exp(.fitted))) + 
  
  # average (needs to be transformed 4 to the right, so that plotting works)
  geom_function(fun = function(x) {exp( mean(coef(lm_log_lin_l$nebraska), na.rm = TRUE)* (x - 5))},
                mapping = aes(linetype = "Mean Decay Rate"),
                color = "blue", size = 1, #linetype = "dashed", 
                inherit.aes = FALSE) + 
  scale_linetype(name = "")


# save the plot
png(filename = paste0(p_output, "plots/nebraska_log2_log_lin.png"), 
    width = 10, height = 5, units = "in", res = 400)
print(plot_grid(gg_nebraska_log2, gg_nebraska_log_lin + theme(legend.position = "none"),
                rel_widths = c(1.4, 1)))
dev.off()


```



```{r basic-persistence-plots}


gg_persist_base <- ggplot(data = dat) + 
  theme_classic() + 
  labs(title = "Persistence of Abandoned Land",
       subtitle = site_df$description[indx],
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + theme(legend.position = "bottom") +
  theme(legend.text = element_text(angle = 320, vjust = 1, hjust = 0))

# raw area
gg_persistence_count <- gg_persist_base + 
  geom_line(mapping = aes(x = age, y = area_ha / 10^3,
                          group = year_abn, color = year_abn), 
            size = 1.25) + 
  labs(y = expression("Area abandoned (10"^{3}*" ha)") , 
       x = "Years since initial abandonment")
  
# as percentage
gg_persistence_proportion <- gg_persist_base + 
  geom_line(mapping = aes(x = age, y = proportion,
                          group = year_abn, color = year_abn), 
            size = 1.25) + 
  labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment")

# na_first ----------- #
gg_persistence_count_na_first <- gg_persist_base + 
  geom_line(mapping = aes(x = year, y = area_ha / 10^3,
                          group = year_abn, color = year_abn), 
            size = 1.25) + 
  labs(y = expression("Area abandoned (10"^{3}*" ha)") , 
       x = "Year")

gg_persistence_proportion_na_first <- gg_persist_base + 
  geom_line(mapping = aes(x = year, y = proportion,
                          group = year_abn, color = year_abn), 
            size = 1.25) + 
  labs(y = "Proportion remaining abandoned", 
       x = "Year")

```


```{r abandonment-persistence}

# average proportion of abandoned land remaining, across time abandoned

df_av <- persistence_list_b1_s$na_last %>%
  mutate(year_abn = as_factor(year_abn)) %>%
  group_by(time_abn) %>%
  summarise(count = n(),
            sd = sd(proportion, na.rm = TRUE), 
            se = sd/sqrt(count),
            proportion = mean(proportion), 
            year_abn = as_factor("mean"))
df_av


# as percentage
gg_persistence_proportion # see basic-persistence-plots chunk

gg_persistence_proportion + geom_line(data = df_av, mapping = aes(x = time_abn, y = proportion),
              size = 3)


# ----------------------------------- plot ------------------------------------------------- #


gg_half_life_s <- gg_persistence_proportion + 
  geom_hline(yintercept = 0.5, linetype="dashed", 
             color = "red", size = 0.75) +
  geom_ribbon(data = df_av, aes(x = time_abn, ymin = proportion - 1.96*se, ymax=proportion + 1.96*se), 
              color = "gray", fill = "gray", alpha=0.5) + 
  geom_line(data = df_av, mapping = aes(x = time_abn, y = proportion),
              size = 2, color = "black")

gg_half_life_s
df_av %>% filter(proportion <= 0.51)



# save basic half-life plot with 
png(filename = paste0(p_output, "plots/half_life", site_label, ".png"), 
    width = 7, height = 5, units = "in", res = 400)
print(gg_half_life_s)
dev.off()
```

```{r exp-geom_smooth}
gg_persist_mod_base <- ggplot(data = dat, mapping = aes(x = age, y = proportion)) + 
  theme_classic() + 
  geom_point(aes(group = year_abn, color = year_abn), size = 1.25) + 
  labs(title = "Persistence of Abandoned Land",
       subtitle = site_df$description[indx],
       color = "Year Abandoned",
       y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment") + 
  scale_color_distiller(palette = "Greens") + 
  theme(legend.position = "bottom", 
        legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", size = 0.75)


# ----------------------------------- plot average proportion ------------------------------------------------- #
gg_persist_mod_base + 
  geom_hline(yintercept = 0.5, linetype="dashed", 
             color = "red", size = 0.75) +
  geom_ribbon(data = df_av, aes(x = time_abn, ymin = proportion - 1.96*se, ymax=proportion + 1.96*se), 
              color = "gray", fill = "gray", alpha=0.5) + 
  geom_line(data = df_av, mapping = aes(x = time_abn, y = proportion),
              size = 2, color = "black")


# ----------------------------------- plot regular lm ------------------------------------------------- #

# regular lm
gg_persist_mod_base + 
  geom_smooth(method = "lm", se = FALSE)


# ----------------------------------- y ~ log(x) ------------------------------------------------- #

gg_persist_mod_base + 
  geom_smooth(method = "lm", formula = y ~ log(x), se = FALSE) + 
  geom_vline(xintercept = 
               exp(( 0.5 - coef(lm(proportion ~ log(age), data = dat))[1] ) / 
                     coef(lm(proportion ~ log(age), data = dat))[2]))

# others
gg_persist_mod_base + 
  geom_smooth(method = "lm", formula = y ~ exp(x), se = FALSE)


# ----------------------------------- log(y) ~ x ------------------------------------------------- #

gg_persist_mod_base + 
  geom_smooth(method = "lm", formula = log(y) ~ x,
              se = FALSE)

# simple lm ----------------- #
lm_mod <- lm(log(proportion) ~ age, data = dat)

  # note, needs to be log transformed
ggplot(data = dat, mapping = aes(x = age, y = log(proportion))) + 
  theme_classic() + 
  geom_point(mapping = aes(group = year_abn, color = year_abn), 
            size = 1.25) + 
  labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land",
       subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + 
  theme(legend.position = "bottom", legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 

  # log(y) ~ x
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
  geom_hline(yintercept = log(0.5), linetype = "dashed", color = "red", size = 0.75) + 
  geom_vline(xintercept = cc_half_life(lm_mod))
```


```{r cc_half_life}
# ----------------------------------- calculate the half-life  ------------------------------------------------- #

augment(lm_mod)
coef(lm_mod)

# log(y) = b0 + b1*x
# log(y) - b0 = b1*x
# ( log(y) - b0 ) / b1 = x
# ( log(0.5) - b0 ) / b1 = half-life
# ( log(0.5) - intercept ) / slope = half-life

(log(0.5) - coef(lm_mod)["(Intercept)"]) / coef(lm_mod)["age"] # = x 

# function to do this, set up to work for both lm and lmer outputs

cc_half_life <- function(model_out, cohort = TRUE, confint = FALSE, 
                         type = "log-log", fixed_effect = "time") {
  
  if(cohort) {
    mean_coef <- tidy(model_out, conf.int = TRUE) %>%
      summarise_at(c("estimate", "conf.low", "conf.high"), mean, na.rm = TRUE)
    
    
    if(type == "log-log") {
      warning("Producing time to 50%, not true half-life. (Log-log relationships do not have true half-lives)")
      half_life_df <- data.frame("half-life" = 0.5^(1/mean_coef$estimate),
                 "conf.low" = 0.5^(1/mean_coef$conf.low),
                 "conf.high" = 0.5^(1/mean_coef$conf.high)
               )
    }
    
    if(type == "semi-log") {
      half_life_df <- data.frame("half-life" = log(0.5)/mean_coef$estimate,
                 "conf.low" = log(0.5)/mean_coef$conf.low,
                 "conf.high" = log(0.5)/mean_coef$conf.high
               )
    }
    
    } else {
      
      params <- tidy(model_out, conf.int = TRUE) %>% as.data.frame()
      
      int <- params[which(params$term == "(Intercept)"), "estimate"]
      int.low <- params[which(params$term == "(Intercept)"), "conf.low"]
      int.high <- params[which(params$term == "(Intercept)"), "conf.high"]
      slope <- params[which(params$term == fixed_effect), "estimate"]
      slope.low <- params[which(params$term == fixed_effect), "conf.low"]
      slope.high <- params[which(params$term == fixed_effect), "conf.high"]
      
      if("(Intercept)" %in% params$term) {
        half_life_df <- data.frame("half-life" = (log(0.5) - int) / slope,
                                   "conf.low" = (log(0.5) - int.low) / slope.low,
                                   "conf.high" = (log(0.5) - int.high) / slope.high
                                   )
        } else {
        half_life_df <- data.frame("half-life" = (log(0.5)) / slope,
                                   "conf.low" = (log(0.5)) / slope.low,
                                   "conf.high" = (log(0.5)) / slope.high
        )
        }
    }
  
  if(fixed_effect == "time") {half_life_df <- half_life_df + 5}
  
  if (confint) {half_life_df} else {half_life_df[, 1]}

}

tidy(lm_mod)
tidy(lmer_mod)

cc_half_life(lm_mod, cohort = FALSE, confint = TRUE, fixed_effect = "time")
cc_half_life(lm_mod, confint = TRUE)
cc_half_life(lmer_mod, cohort = FALSE, confint = TRUE, fixed_effect = "age")

cc_half_life(lmer_mod_slope, fixed_effect = "age", cohort = FALSE, confint = TRUE)
cc_half_life(lmer_mod_slope0, fixed_effect = "time", cohort = FALSE, confint = TRUE)


tidy(lm_mod_log2_cohort) %>% select(estimate) %>% unlist() %>% mean(., na.rm = TRUE)



cc_half_life(lm_mod_log2_cohort, type = "log-log", cohort = TRUE, confint = TRUE)
cc_half_life(lm_mod_cohort_slope0, type = "semi-log", cohort = TRUE, confint = TRUE)

cc_half_life(lm_mod_log2_cohort, type = "log-log")
cc_half_life(lm_mod_log2_cohort)
cc_half_life(lm_mod_cohort_slope0, type = "semi-log")


mean_coef$conf.high

0.5^(1/mean_coef$estimate)

cc_half_life(lm_mod_log2_cohort, cohort = TRUE, type = "log-log", confint = TRUE)


```



```{r lm}
# as shown above, the lm call to model basic exponential decay is as follows:
mod_decay_s <- lm(log(proportion) ~ age, data = dat) # or
lm_mod <- lm(log(proportion) ~ time, data = dat)


# extract coefficients, summary, confint, etc., along with other useful functions
summary(lm_mod)
coef(lm_mod)
confint(lm_mod)
tidy(lm_mod, conf.int = TRUE) 

augment(lm_mod) # to get predicted values, etc. along with the input response and predictor variables

cc_half_life(lm_mod, confint = TRUE)

# ------- plot half life with lm curve ------------------------------------------------- #

# two ways to plot the model results, either
gg_persist_mod_base + 
  
  # add fitted values, exponentiated
  geom_line(data = augment(lm_mod, newdata = mutate(data.frame(age = 5:30), time = age - 5)), 
            aes(x = age, y = exp(.fitted)),
            color = "red", size = 2) + 
  
  # or add as a function
  geom_function(fun = function(x) {exp(coef(lm_mod)[2]*x + coef(lm_mod)[1])},
                color = "blue", size = 1)




# fancier plot
# prepare to be plotted
model_estimates <- data.frame(
  age = 5:30,
  int = tidy(lm_mod) %>% filter(term == "(Intercept)") %>% select(estimate) %>% unlist(use.names = FALSE),
  int_low = tidy(lm_mod, conf.int = TRUE) %>% filter(term == "(Intercept)") %>% select(conf.low) %>% unlist(use.names = FALSE),
  int_high = tidy(lm_mod, conf.int = TRUE) %>% filter(term == "(Intercept)") %>% select(conf.high) %>% unlist(use.names = FALSE),
  slope = tidy(lm_mod) %>% filter(term == "age") %>% select(estimate) %>% unlist(use.names = FALSE),
  slope_low = tidy(lm_mod, conf.int = TRUE) %>% filter(term == "age") %>% select(conf.low) %>% unlist(use.names = FALSE),
  slope_high = tidy(lm_mod, conf.int = TRUE) %>% filter(term == "age") %>% select(conf.high) %>% unlist(use.names = FALSE)
  ) %>%
  mutate(estimate = exp(int + slope* age),
         lower = exp(int_low + slope_low* age),
         upper = exp(int_high + slope_high* age))


# much easier to do this:

gg_half_life_lm_s <- gg_persist_mod_base + 
  # make and add confidence interval
  geom_ribbon(data = mutate(data.frame(age = 5:30),
                            lower = exp(confint(lm_mod)[2, 1]*age +
                                          confint(lm_mod)[1, 1]),
                            upper = exp(confint(lm_mod)[2, 2]*age +
                                          confint(lm_mod)[1, 2])),
              mapping = aes(ymin = lower, ymax = upper, x = age), inherit.aes = FALSE, 
              fill = "dark gray", alpha = 0.5, color = "dark gray", linetype = "solid") +
  geom_function(fun = function(x) {exp(coef(lm_mod)[2]*x + 
                                         coef(lm_mod)[1])},
                color = "dark blue", size = 1) + 
  annotate("text", x = 24, y = 0.8, label = "half-life = 23.17 years", fontface = "italic", angle = 90) + 
  geom_vline(xintercept = cc_half_life(lm_mod))




# save
png(filename = paste0(p_output, "plots/half_life_lm_s.png"), 
    width = 5, height = 5, units = "in", res = 400)
print(gg_half_life_lm_s)
dev.off()






# with coord_trans
ggplot(data = dat %>% mutate(year_abn = as.numeric(levels(year_abn))[year_abn]),
       mapping = aes(x = (age), y = (proportion), group = year_abn, color = year_abn)) + 
  # theme_classic() + 
  geom_point(size = 1) + 
  labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment") + 
  scale_color_distiller(palette = "Greens") +
  coord_trans(y = "log", x = "log") 
#+ scale_y_log10()
```

```{r lm0-int}

# -------------------- consistent intercept ----------------------------- #
# dat_shift

dat$time # this has been rescaled to allow for the intercept at (0, 1)

lm_mod0 <- lm(log(proportion) ~ 0 + time, data = dat)
# note, this is the same as running: lm(log(proportion) ~ 0 + I(age - 5), data = dat)

lm_mod0
coef(lm_mod0)
# y = exp(beta1 * t)
exp( coef(lm_mod0) * 0)
for (t in 0:10) {
  print(
    exp( coef(lm_mod0) * t)
  )
}
# each time step, the amount from the previous time step is multiplied by e ^ coef
exp(coef(lm_mod0))

exp( coef(lm_mod0) * 5) / 
  exp( coef(lm_mod0) * 4)



lm(log(proportion) ~ time, data = dat) # note, this is the same as lm(formula = log(proportion) ~ age, data = dat) (lm_mod)
lm(log(proportion) ~ 0 + age, data = dat) # this doesn't quite work, since it sets the intercept to (0, 1), since the proportions start at age = 5. 
lm_mod0

# Note that lm(y ~ x, data = dat) is the same as lm(y ~ 1 + x, data = dat). 
# The 1 in the longer notation represents that we want lm to estimate an intercept. 
# If we change this to a 0, explicitly, lm only calculates a slope coefficient for x.  This is what we've done above. 


# lm_mod_offset <- lm(log(proportion) ~ 0 + time, data = dat)
# lm_mod_offset1 <- lm(log(proportion) ~ time, data = dat)


ggplot(data = dat, mapping = aes(x = time, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  
  # lm_mod0
  geom_line(data = augment(lm_mod0, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = time, y = exp(.fitted)),
            color = "red", size = 1, inherit.aes = FALSE)


# 0 + age
ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ 0 + age") + 
  geom_function(fun = function(x) {exp( coef(lm(log(proportion) ~ 0 + age, data = dat)) * (x - 5))},
                color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))

coef(lm(log(proportion) ~ 0 + age, data = dat))


# comparing to the original lm_mod
gg_persist_mod_base + 
  geom_line(data = augment(lm_mod, newdata = data.frame(age = 5:30)), 
            aes(x = age, y = exp(.fitted)), 
            color = "red", size = 2)

gg_persist_mod_base + scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5)) + 
  
  # lm_mod
  geom_line(data = augment(lm_mod, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = "green", size = 1, inherit.aes = FALSE) + 
  
  # lm_mod0
  geom_line(data = augment(lm_mod0, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = "red", size = 1, inherit.aes = FALSE) + 
  
  # age, slope only, starting at 1. 
  geom_line(data = augment(lm(log(proportion) ~ 0 + age, data = dat), 
                           newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = "blue", size = 1, inherit.aes = FALSE)
  



# or...
ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() +
  geom_line(data = augment(lm_mod, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = "green", size = 1, inherit.aes = FALSE) + 
  
  # add fitted values, exponentiated
  geom_line(data = augment(lm_mod0, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = "red", size = 1, inherit.aes = FALSE) + 
  scale_color_distiller(palette = "Greens")





```


```{r lm-log2}
# log2

lm_mod_log2 <- lm(log(proportion) ~ log(age), data = dat)
summary(lm_mod)
summary(lm_mod_cohort)
summary(lm_mod_log2)


# log-log plot
ggplot(data = dat, mapping = aes(x = log(age), y = log(proportion), group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens")

ggplot(data = dat, mapping = aes(x = log(time_1), y = log(proportion), group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_smooth(method = "lm", se = 0)



plot_grid(ggplot(data = dat, mapping = aes(x = log(age), y = log(proportion), group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_smooth(method = "lm", se = 0, formula = y ~ x),
  ggplot(data = dat, mapping = aes(x = log(time_1), y = log(proportion), group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_smooth(method = "lm", se = 0, formula = y ~ 0 + x))


ggplot(data = dat, 
       mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() +
  scale_color_distiller(palette = "Greens") + 
  coord_trans(y = "log", x = "log") + 
  geom_smooth(method = "lm", se = 0)




ggplot(data = dat, mapping = aes(x = log(age), y = log(proportion), group = year_abn, color = year_abn)) + 
  geom_point() +
  # add fitted values, exponentiated
  # geom_line(data = augment(lm_mod, newdata = data.frame(age = 5:30)),
  #           aes(x = age, y = exp(.fitted)),
  #           color = "red", size = 1, inherit.aes = FALSE) +
  # geom_line(data = mutate(augment(lm_mod_cohort), year_abn = as.numeric(levels(cohort))[cohort]),
  #           aes(x = age, y = exp(.fitted))) + 
  scale_color_distiller(palette = "Greens") + 
  geom_smooth(method = "lm", se = 0, aes(x = log(age), y = log(proportion)), inherit.aes = FALSE)



ggplot(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  theme_classic() + 
  labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land: Shaanxi Province",
       subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + 
  theme(legend.position = "bottom",
        legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  scale_x_continuous(n.breaks = 10) +
  
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", size = 0.75) + 
  geom_point(data = dat) +
  
  geom_line(data = augment(lm_mod_log2, newdata = data.frame(age = 5:30)),
            aes(x = age, y = exp(.fitted)),
            color = gg_colors[5], size = 1.5, inherit.aes = FALSE) + 
  geom_vline(xintercept = (0.5/exp(coef(lm_mod_log2)[1]))^(1/coef(lm_mod_log2)[2]), 
             color = gg_colors[5]) + 
  annotate("text", fontface = "italic", angle = 90, size = 3.5,
           x = cc_half_life(lm_mod_log2, fixed_effect = "time") + 5, 
           y = 0.85, label = paste0(
             "lm_mod0: \n", 
             round((0.5/exp(coef(lm_mod_log2)[1]))^(1/coef(lm_mod_log2)[2]), 2), 
             " years")) + 
  geom_function(fun = function(x) { x^(coef(lm_mod_log2)[2])*exp(coef(lm_mod_log2)[1])},
                color = gg_colors[5], size = 1, linetype = "dashed", inherit.aes = FALSE)
  


# log - log, with intercept 0
lm_mod_log2_0 <- lm(log(proportion) ~ 0 + log(time_1), data = dat)

ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ 0 + log(time_1):cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1), data = dat))),
            aes(x = exp(`log(time_1)`) + 4, y = exp(.fitted)), inherit.aes = FALSE) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))



```



```{r lm-cohort}

# add cohort as a fixed effect predictor variable. This is analogous to a parallel slopes lm, where cohort just changes the intercept coefficient, and age has the same role across all

lm_mod
lm_mod_cohort_int <- lm(log(proportion) ~ time + cohort - 1, data = dat) # parallel slopes, unique intercepts
lm_mod_cohort_slope <- lm(log(proportion) ~ time + time:cohort, data = dat) # single intercept, unique slopes

lm_mod_cohort_slope0 <- lm(log(proportion) ~ 0 + time:cohort, data = dat) # no global intercept (i.e. 1), unique slopes

lm_mod_cohort_x <- lm(log(proportion) ~ time * cohort - 1, data = dat) # allowing for all cohorts to have individual intercepts and slopes with age

lm_mod_cohort

coef(lm_mod_cohort)
summary(lm_mod_cohort)
tidy(lm_mod_cohort, conf.int = TRUE) 
augment(lm_mod_cohort)
coef(lm_mod_cohort)


ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 

  # lm_mod_cohort_slope0
  geom_line(data = mutate(augment(lm_mod_cohort_slope0), 
                          year_abn = as.numeric(levels(cohort))[cohort], 
                          age = time + 5),
            aes(x = age, y = exp(.fitted))) + 
  
  # geom_function(fun = function(x) {exp( mean(coef(lm_mod_cohort_slope0), na.rm = TRUE)*x)},
  #               color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE)
  
  geom_function(fun = function(x) {exp( mean(coef(lm_mod_cohort_slope0), na.rm = TRUE)* (x - 5))},
                color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE)


# plot
ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 

  # lm_mod
  # geom_line(data = augment(lm_mod, newdata = data.frame(age = 5:30)),
  #           aes(x = age, y = exp(.fitted)),
  #           color = "red", size = 1, inherit.aes = FALSE) +

  # lm_mod_cohort_slope0
  geom_line(data = mutate(augment(lm_mod_cohort_slope0), 
                          year_abn = as.numeric(levels(cohort))[cohort], 
                          age = time + 5),
            aes(x = age, y = exp(.fitted))) + 
  
  # geom_function(fun = function(x) {exp( mean(coef(lm_mod_cohort_slope0), na.rm = TRUE)*x)},
  #               color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE)
  
  geom_function(fun = function(x) {exp( mean(coef(lm_mod_cohort_slope0), na.rm = TRUE)* (x - 5))},
                color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE) + 
  geom_vline(xintercept = 26.26429)

  
  # add individual cohort curves
  # geom_function(fun = function(x) {exp(coef(lm_mod_cohort)["age"]*x + 
  #                                       coef(lm_mod_cohort)["cohort1988"])},
  #               color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE) +
  # geom_function(fun = function(x) {exp(coef(lm_mod_cohort)["age"]*x + 
  #                                       coef(lm_mod_cohort)["cohort2010"])},
  #               color = "orange", size = 1, linetype = "solid", inherit.aes = FALSE)
  



          
p1 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  
  labs(title = "log(proportion) ~ time + cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ time + cohort, data = dat)),
                          year_abn = as.numeric(levels(cohort))[cohort], age = time + 5),
            aes(x = age, y = exp(.fitted)))

p2 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  
  labs(title = "log(proportion) ~ time + time:cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ time + time:cohort, data = dat)),
                          year_abn = as.numeric(levels(cohort))[cohort], age = time + 5),
            aes(x = age, y = exp(.fitted)))

p3 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 

  labs(title = "log(proportion) ~ cohort + time:cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ cohort + time:cohort, data = dat)),
                          year_abn = as.numeric(levels(cohort))[cohort], age = time + 5),
            aes(x = age, y = exp(.fitted)))

p4 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 

  labs(title = "log(proportion) ~ time + cohort + time:cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ time + cohort + time:cohort, data = dat)), 
                          year_abn = as.numeric(levels(cohort))[cohort], age = time + 5),
            aes(x = age, y = exp(.fitted)))

p5 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 

  labs(title = "log(proportion) ~ 0 + time:cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + time:cohort, data = dat)), 
                          year_abn = as.numeric(levels(cohort))[cohort], age = time + 5),
            aes(x = age, y = exp(.fitted)))
lm_mod_cohort_slope0

p6 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 

  labs(title = "log(proportion) ~ 0 + log(time_1):cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1):cohort, 
                                     data = mutate(dat, time_1 = time + 1))), 
                          year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted))) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))



lm(log(I(proportion - 1)) ~ log(I(age - 5)):cohort, data = dat)

lm_mod_log2_cohort

dat %>% mutate(time2 = time + 1, 
                log_time2 = log(time2))

lm(log(I(proportion - 1)) ~ log(I(age - 5)):cohort, data = dat)

# newdata = data.frame(age = 5:30, cohort = factor(rep(1988:2013))
# geom_function(fun = function(x) { x^(coef(lm_mod_log2)[2])*exp(coef(lm_mod_log2)[1])},
#                 color = gg_colors[5], size = 1, linetype = "dashed", inherit.aes = FALSE)



# plot_grid
plot_grid(p1, p2, p3, p4, p5, p6)
```

```{r lm-log2-cohort}
# lm_mod_log2_cohort <- lm(log(proportion) ~ 0 + log(age):cohort, data = dat)

lm_mod_log2_cohort <- lm(log(proportion) ~ 0 + log(time_1):cohort, data = dat)
summary(lm_mod_log2_cohort)
tidy(lm_mod_log2_cohort)
glance(lm_mod_log2_cohort)
formula(lm_mod_log2_cohort)
augment(lm_mod_log2_cohort)

# compare time_1 vs. age:
lm(log(proportion) ~ log(time_1) + cohort + log(time_1):cohort, data = dat)
lm(log(proportion) ~ log(time_1):cohort, data = dat)
lm(log(proportion) ~ log(age):cohort, data = dat)
lm(log(proportion) ~ log(time_1):cohort, data = dat)
lm(log(proportion) ~ log(age):cohort, data = dat)



# plot time_1
ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_line(data = mutate(augment(lm_mod_log2_cohort), 
                          time_1 = exp(`log(time_1)`),
                          proportion = exp(.fitted),
                          year_abn = as.numeric(levels(cohort))[cohort])) + 
  scale_x_continuous(n.breaks = 30) + 
  geom_function(fun = function(x) { x ^ coef(lm_mod_log2_cohort)[1]},
                size = 1, linetype = "dashed", inherit.aes = FALSE)


# plot with age
ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  #scale_x_continuous(breaks = seq(from = 0, to = 40, by = 5), limits = c(5, 40)) + 
  geom_hline(yintercept = 0.5) + 
  labs(subtitle = "log(proportion) ~ 0 + log(time_1):cohort") + 
  geom_line(data = mutate(augment(lm_mod_log2_cohort), 
                          year_abn = as.numeric(levels(cohort))[cohort]),
            aes(x = exp(`log(time_1)`) - 1 + 5, y = exp(.fitted))) + 
  
  # average (needs to be transformed 4 to the right, so that plotting works)
  geom_function(fun = function(x) { (x - 4) ^ (mean(coef(lm_mod_log2_cohort), na.rm = TRUE))},
                color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE) +
  
  geom_line(data = mutate(augment()))
  
  # half-life: cc_half_life(lm_mod_log2_cohort)
  geom_vline(xintercept = cc_half_life(lm_mod_log2_cohort))


# log log scale
ggplot(data = dat, mapping = aes(x = log(time_1), y = log(proportion), group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  # geom_smooth(method = "lm", se = 0) + 
  # geom_line(data = mutate(augment(lm_mod_log2_cohort), 
  #                         time_1 = exp(`log(time_1)`),
  #                         proportion = exp(.fitted),
  #                         year_abn = as.numeric(levels(cohort))[cohort]))#,
  #           #aes(x = exp(`log(time_1)`) - 1 + 5, y = exp(.fitted)))
  
  # alternatively
  geom_line(data = mutate(augment(lm_mod_log2_cohort), 
                          time_1 = exp(`log(time_1)`),
                          proportion = exp(.fitted),
                          year_abn = as.numeric(levels(cohort))[cohort]),
            aes(x = `log(time_1)`, y = .fitted))





1 ^ mean(coef(lm_mod_log2_cohort), na.rm = TRUE)
2 ^ mean(coef(lm_mod_log2_cohort), na.rm = TRUE)






# comparison plots:

p1 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  
  labs(title = "log(proportion) ~ time + cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ time + cohort, data = dat)),
                          year_abn = as.numeric(levels(cohort))[cohort], age = time + 5),
            aes(x = age, y = exp(.fitted)))

p7 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  
  labs(title = "log(proportion) ~ time + cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1), data = dat)),
                          # year_abn = as.numeric(levels(cohort))[cohort], 
                          age = time + 5),
            aes(x = age, y = exp(.fitted)))



# is the mean coefficient the same as lm without cohort?

pm1 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  
  labs(title = "log(proportion) ~ 0 + log(time_1)") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1), 
                                     data = filter(dat_l, site == "shaanxi"))),
                          # year_abn = as.numeric(levels(cohort))[cohort], 
                          time_1 = exp(`log(time_1)`),
                          age = time_1 + 4),
            aes(x = age, y = exp(.fitted)), size = 1, inherit.aes = FALSE)

mean_coef_df[9, 2] %>% as.numeric()

mean(coef(lm_log2_l$shaanxi), na.rm = TRUE)

lm_log2_l$shaanxi
lm(log(proportion) ~ 0 + log(time_1), 
   data = filter(dat_l, site == "shaanxi"))


plot(lm(log(proportion) ~ 0 + log(time_1), 
        data = filter(dat_l, site == "shaanxi")))

glance(lm(log(proportion) ~ 0 + log(time_1), 
          data = filter(dat_l, site == "shaanxi")))

w_mean <- tidy(lm_log2_l$shaanxi) %>% 
  mutate(weight = 26:1) %>% 
  filter(!is.na(estimate)) %>%
  group_by() %>% 
  summarise(weighted_mean = sum(estimate * weight) / sum(weight)) %>%
  as.numeric()

lm(log(proportion) ~ 0 + log(time_1), 
          data = filter(dat_l, site == "shaanxi"))

dat %>% group_by(cohort) %>% summarise(n = n()) %>% print(n = 30)

# save
png(filename = paste0(p_output, "plots/mean_coef_alt.png"), 
    width = 6, height = 5, units = "in", res = 400)

pl1 + 
  geom_function(fun = function(x) { (x - 4) ^ mean(coef(lm_mod_log2_cohort), na.rm = TRUE)},
                size = 1, color = "red", linetype = "dashed", inherit.aes = FALSE) + 
  geom_function(fun = function(x) { (x - 4) ^ w_mean},
                size = 1, color = "red", inherit.aes = FALSE) + 
  geom_function(fun = function(x) { (x - 4) ^ coef(lm(log(proportion) ~ 0 + log(time_1), 
                                                      data = filter(dat_l, site == "shaanxi")))},
                size = 1, color = "blue", inherit.aes = FALSE) + 
  labs(title = "Alternative methods for calculating site mean decay coefficient") +
  annotate("text", fontface = "bold", x = 25, y = 0.68, label = "mean coef", size = 3.5, angle = -12) + 
  annotate("text", fontface = "bold", x = 26, y = 0.60, label = "weighted mean", size = 3.5, angle = -12) + 
  annotate("text", fontface = "bold", x = 27, y = 0.56, label = "lm w/o cohort f.e.", size = 3.5, angle = -12)

dev.off()

```

# Intuition behind log-log
An exponential decay curve (lon-linear, or semi-log) makes intuitive sense, if we assume that the amount of abandoned land declines by a constant proportion each time step. In this case, the amount at t+1 = the amount at t multiplied by e^(coef) . The log-lin relationship is as follows, lm( log(y) ~ x ):
log(y) = b1 * x + C
assuming no intercept
log(y) = b1 * x
e ^ log(y) =  e ^ (b1 * x)
y = e ^ (b1 * x)


A log-log curve corresponds to a monomial ( y = a * x^k ), where the slope of the log-log line corresponds to k, the power term.
lm( log(y) ~ log(x) ):
log(y) = b1 * log(x) + C
assuming no intercept
log(y) = b1 * log(x)
e ^ log(y) = e ^ ( b1 * log(x) )
y = ( e ^ log(x) ) ^ (b1)
y = x ^ b1

With multiple cohorts, the math works out like this
log(y) = b1 * log(x) * cohort1 + b2 * log(x) * cohort2 + b3 * log(x) * cohort3 + ... + bn * log(x) * cohortn; where n is the number of cohorts (in our case, 26)

e ^ log(y) = e ^ ( b1 * log(x) * cohort1 + b2 * log(x) * cohort2 + b3 * log(x) * cohort3 + ... )
y = e ^ (b1 * log(x) * cohort1) * e ^ (b2 * log(x) * cohort2) * e ^ (b3 * log(x) * cohort3) * e ^ (...)

cohort1, cohort2, etc. are dummy variables that are 1 or 0, depending on whether it is that cohort or not. So, all the other terms reduce to multiplying by e ^ 0 or 1. So, each cohort's curve reduces to:
y = x ^ b1

This means that rather than declining by a constant proportion each time step, as with the exponential decay function (e ^ coef), the decay rate changes as a function of time. The proportional decrease over time is equal to:
y(x+1) / y(x)
(x + 1) ^ b1 / x ^ b1
((x + 1) / x ) ^ b1

So, from t = 4 to t = 5, the amount of abandoned land declines by (5 / 4) ^ b1 
As t gets large, ((t + 1) / t) gets closer and closer to 1. 

```{r log-intuition}
ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_line(data = mutate(augment(lm_mod_log2_cohort), 
                          time_1 = exp(`log(time_1)`),
                          proportion = exp(.fitted),
                          year_abn = as.numeric(levels(cohort))[cohort])) + 
  scale_x_continuous(n.breaks = 30) + 
  geom_function(fun = function(x) { x ^ coef(lm_mod_log2_cohort)[1]},
                size = 1, linetype = "dashed", inherit.aes = FALSE)
  


# cohort 1's function:
# y = x ^ b1
coef(lm_mod_log2_cohort)[1]
# y = 1 ^ ()

mutate(augment(lm_mod_log2_cohort), 
       time_1 = exp(`log(time_1)`),
       proportion = exp(.fitted))




ft <- augment(lm_mod_log2_cohort, 
              newdata = data.frame(time_1 = 1:26, 
                                   cohort = factor(1988))) %>% 
  mutate(y = exp(.fitted))

for (i in 1:26) {print(ft$y[i + 1] / ft$y[i])}

(2/1)^coef(lm_mod_log2_cohort)[1]
(5/4)^coef(lm_mod_log2_cohort)[1]



1 ^ coef(lm_mod_log2_cohort)[1]
2 ^ coef(lm_mod_log2_cohort)[1]
3 ^ coef(lm_mod_log2_cohort)[1]
4 ^ coef(lm_mod_log2_cohort)[1]
5 ^ coef(lm_mod_log2_cohort)[1]

data.frame(time = 1:26) %>% mutate(y = time ^ coef(lm_mod_log2_cohort)[1]) %>% head()



```


```{r cohort-v-year_abn}
# ---- #
# comparing cohort as a categorical variable to year_abn as a continuous variable.
pl0 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ 0 + log(time_1)") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1), data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5
                          ),
            aes(x = age, y = exp(.fitted)), inherit.aes = FALSE) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))

pl1 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ 0 + log(time_1):cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1):cohort, 
                                     data = mutate(dat, time_1 = time + 1))), 
                          year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted))) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))

pl2 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ 0 + log(time_1):year_abn") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1):year_abn, 
                                     data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted), size = year_abn)) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))

pl3 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ 0 + log(time_1)*year_abn") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1)*year_abn, 
                                     data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))


# plot_grid
plot_grid(pl0, pl1, pl2, pl3)

lm(log(proportion) ~ 0 + log(time_1)*year_abn, data = dat) %>% coef()
test1 <- lm(log(proportion) ~ 0 + log(time_1) + year_abn + log(time_1):year_abn, data = dat)
test2 <- lm(log(proportion) ~ 0 + log(time_1) + log(time_1):year_abn, data = dat)
lm(log(proportion) ~ 0 + year_abn + log(time_1):year_abn, data = dat) %>% coef()

coef(test1)
coef(test2)

coef(test1)[1] + coef(test1)[3] * 1988
coef(test2)[1] + coef(test2)[2] * 1988






pl4 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(subtitle = "log(proportion) ~ 0 + log(time_1) + year_abn + log(time_1):year_abn") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1) + year_abn + log(time_1):year_abn, 
                                     data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))


# this one doesn't make sense
pl5 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(subtitle = "log(proportion) ~ 0 + year_abn + log(time_1):year_abn") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + year_abn + log(time_1):year_abn, 
                                     data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))

# ** if taking the year_abn approach, this specification should be used.
pl6 <- ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(subtitle = "log(proportion) ~ 0 + log(time_1) + log(time_1):year_abn") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1) + log(time_1):year_abn, data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))



# comparing the two
ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(subtitle = "log(proportion) ~ 0 + log(time_1) + year_abn + log(time_1):year_abn") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1) + year_abn + log(time_1):year_abn, 
                                     data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1, color = "red") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1) + log(time_1):year_abn, 
                                     data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1, color = "blue") + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))


# plot_grid
plot_grid(pl3, pl4, pl5, pl6)




# compare AIC
lm_mod_log2_cohort
lm_mod_log2_year_abn <- lm(log(proportion) ~ 0 + log(time_1) + log(time_1):year_abn, data = dat)

glance(lm_mod_log2_cohort) %>% select(AIC)
glance(lm_mod_log2_year_abn) %>% select(AIC) 
# conclusion: allowing the coefficients to vary between cohorts as categorical variables, rather than as continuous variables, improves the model fit dramatically.


lm(log(proportion) ~ 0 + log(time_1) + log(time_1):year_abn, data = dat)


ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  # lm_mod_log2_cohort
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1):cohort, 
                                     data = dat)), 
                          year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1, color = "red") + 
  
  # with year_abn, even spacing between cohorts.
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1) + log(time_1):year_abn, 
                                     data = dat)), 
                          #year_abn = as.numeric(levels(cohort))[cohort],
                          age = exp(`log(time_1)`) - 1 + 5),
            aes(x = age, y = exp(.fitted)), size = 1, color = "blue") + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))

```

```{r residual-plots}
lm_mod_log2_cohort

plot(lm_mod_log2_cohort)
augment(lm_mod_log2_cohort) %>% arrange(desc(.hat))

resid_df <- bind_cols(select(dat, year, year_abn, initial_area_abn, 
                             area_ha, age, proportion, time, time_1, 
                             cohort1 = cohort),
                      augment(lm_mod_log2_cohort))

gg_resid_1 <- ggplot(data = resid_df, aes(x = time, y = .std.resid, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + labs(y = "Residuals", x = "Time Past Abandonment Threshold")

gg_resid_2 <- ggplot(data = resid_df, aes(x = initial_area_abn, y = .std.resid, color = year_abn)) + 
  geom_point(position = position_jitter(width = 20)) + scale_color_distiller(palette = "Greens") + 
  labs(y = "Residuals", x = "Initial Area Abandoned") + theme(legend.position = "none")


# save
png(filename = paste0(p_output, "plots/log2_cohort_resid_s.png"), 
    width = 10, height = 5, units = "in", res = 400)
plot_grid(gg_resid_1, gg_resid_2, rel_widths = c(1.2, 1))
dev.off()



```

```{r save-plots}

# ----------------------- log-log cohort plot ----------------------- #

gg_log2_cohort <- ggplot(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point(data = dat) + scale_color_distiller(palette = "Greens") + 
  scale_x_continuous(n.breaks = 30) + 
  geom_hline(yintercept = 0.5, linetype = "dashed") + 
  labs(title = "Decay of abandoned land over time", subtitle = site_df$description[indx], caption = "lm formula: log(proportion) ~ 0 + log(time_1):cohort",
       x = "Time since initial abandonment (years)", y = "Proportion of abandoned land remaining",
       color = "Cohort (Year \nFirst Abandoned)") + 
  geom_line(data = mutate(augment(lm_mod_log2_cohort, newdata = dat0),
                          proportion = exp(.fitted))) + 
  
  # average (needs to be transformed 4 to the right, so that plotting works)
  geom_function(fun = function(x) { (x-4) ^ (mean(coef(lm_mod_log2_cohort), na.rm = TRUE))},
                mapping = aes(linetype = "Mean Decay Rate"),
                color = "blue", size = 1, #linetype = "dashed", 
                inherit.aes = FALSE) + 
  scale_linetype(name = "")


# save the plot
png(filename = paste0(p_output, "plots/log2_cohort_s.png"), 
    width = 8, height = 6, units = "in", res = 400)
print(gg_log2_cohort)
dev.off()



# ----------------------- diagnostics plot ----------------------- #
png(filename = paste0(p_output, "plots/log2_cohort_diag_s.png"), 
    width = 8, height = 8, units = "in", res = 400)
par(mfrow = c(2,2))
plot(lm_mod_log2_cohort)
dev.off()



# ----------------------- log-log cohort residuals plot ----------------------- #


```


```{r add-initial-area-abn}
lm_mod_log2_cohort_area <- lm(log(proportion) ~ 0 + log(time_1):cohort:initial_area_abn, data = dat)

summary(lm_mod_log2_cohort_area)
tidy(lm_mod_log2_cohort_area)
glance(lm_mod_log2_cohort_area)
glance(lm_mod_log2_cohort)

ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ time + cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1):cohort, data = dat)),
                          year_abn = as.numeric(levels(cohort))[cohort], 
                          time_1 = exp(`log(time_1)`),
                          proportion = exp(.fitted)))

ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ time + cohort") + 
  geom_line(data = augment(lm(log(proportion) ~ 0 + log(time_1)*cohort:initial_area_abn, data = dat), 
                           newdata = dat0, interval = "confidence") %>%
              mutate(proportion = exp(.fitted), lower = exp(.lower), upper = exp(.upper)))


# plot time_1
ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_line(data = mutate(augment(lm_mod_log2_cohort), 
                          time_1 = exp(`log(time_1)`),
                          proportion = exp(.fitted),
                          year_abn = as.numeric(levels(cohort))[cohort])) + 
  scale_x_continuous(n.breaks = 30) + 
  geom_function(fun = function(x) { x ^ coef(lm_mod_log2_cohort)[1]},
                size = 1, linetype = "dashed", inherit.aes = FALSE)


ggplot(data = dat, mapping = aes(x = cohort, y = initial_area_abn)) + geom_point()

# compare different specifications:
pa1 <- ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(subtitle = "log(proportion) ~ 0 + log(time_1):cohort:initial_area_abn") + 
  geom_line(data = augment(lm(log(proportion) ~ 0 + log(time_1):cohort:initial_area_abn, data = dat), 
                           newdata = dat0, interval = "confidence") %>%
              mutate(proportion = exp(.fitted), lower = exp(.lower), upper = exp(.upper)))

pa2 <- ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(subtitle = "log(proportion) ~ 0 + log(time_1):initial_area_abn") + 
  geom_line(data = augment(lm(log(proportion) ~ 0 + log(time_1):initial_area_abn, data = dat), 
                           newdata = dat0, interval = "confidence") %>%
              mutate(proportion = exp(.fitted), lower = exp(.lower), upper = exp(.upper)))

pa3 <- ggplot(data = dat, mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(subtitle = "log(proportion) ~ 0 + log(time_1):cohort*initial_area_abn") + 
  geom_line(data = augment(lm(log(proportion) ~ 0 + log(time_1):cohort*initial_area_abn, data = dat), 
                           newdata = dat0, interval = "confidence") %>%
              mutate(proportion = exp(.fitted), lower = exp(.lower), upper = exp(.upper)))



plot_grid(pl1, pa1, pa2, pa3)

augment(lm(log(proportion) ~ 0 + log(time_1):cohort, data = filter(dat, cohort %in% c(1988:2012))), 
        newdata = filter(dat0, cohort %in% c(1988:2012)), interval = "confidence") %>%
  mutate(proportion = exp(.fitted))


dat %>% arrange(desc(year_abn))


ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  # lm_mod_log2_cohort
  geom_line(data = augment(lm(log(proportion) ~ 0 + log(time_1):cohort, data = dat), 
                           newdata = dat0, interval = "confidence") %>%
              mutate(proportion = exp(.fitted)), 
            size = 1, color = "red") + 
  
  # 
  geom_line(data = augment(lm(log(proportion) ~ 0 + log(time_1):initial_area_abn, data = dat), 
                           newdata = dat0, interval = "confidence") %>%
              mutate(proportion = exp(.fitted)), 
            color = "blue") + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))


coef(lm(log(proportion) ~ 0 + log(time_1):cohort + log(time_1):cohort:initial_area_abn, data = dat))
coef(lm(log(proportion) ~ 0 + log(time_1):cohort, data = dat))
coef(lm(log(proportion) ~ 0 + log(time_1):initial_area_abn + log(time_1):cohort, data = dat))

coef(lm(log(proportion) ~ 0 + log(time_1):initial_area_abn, data = dat))


```


```{r lmer}
# Since each cohort of abandoned land represents a particular group, with related values, I will model each cohort as a random effect, with random intercept and random slopes

# here's how the lmer call will work:


# lmer_mod <- lmer(
#   log(proportion) ~  # response variable
#        age + # fixed effect (i.e. what we're curious about)
#        (1 + age|cohort), # random effects (modeling coefficients as random values drawn from a common error distribution, i.e. essentially random differences that we want to control for). 
#      # The 1 represents the random intercept, meaning that the initial starting point can vary for each cohort (year_abn). If this was all we wanted, but with the same relationship to age for each cohort, we'd just have (1|year_abn)
#      # The "age|year_abn" term represents the random slopes, meaning that the relationship between proportion and age (the slope/coefficient) is allowed to vary across the cohorts.
#      data = dat) 

# random slopes, constant intercept

lmer_mod_slope <- lmer(log(proportion) ~ age + (0 + age|cohort), data = dat)
lmer(log(proportion) ~ age + (0 + age|cohort), data = dat) %>% coef()


# random slopes and random intercepts
lmer_mod_slope_int <- lmer(log(proportion) ~ age + (1 + age|cohort), data = dat) # or age + (age|cohort)
lmer(log(proportion) ~ age + (1 + age|cohort), data = dat) %>% coef()


# just random intercepts would look like this
lmer_mod_int <- lmer(log(proportion) ~ age + (1|cohort), data = dat)
lmer(log(proportion) ~ age + (1|cohort), data = dat) %>% coef()




# -------------------------- lmer offset ---------------------------------------- #
# lmer offset: random slopes, but only slope, i.e. no global intercept (so that the intercept is forced to be 1).
# this is the one I intend to use.
lmer_mod_slope0 <- lmer(log(proportion) ~ 0 + time + (0 + time|cohort), data = dat)
# ---------------------------------------------------------------------------- #





lmer_mod_slope
lmer_mod_slope_int
lmer_mod_int
lmer_mod_slope0

augment(lmer_mod_slope0)
fixef(lmer_mod_slope0)*1
exp(fixef(lmer_mod_slope0)*0)

coef(lmer_mod_slope0)
ranef(lmer_mod_slope0)$cohort
fixef(lmer_mod_slope0)

-0.03543644 + ranef(lmer_mod_slope0)$cohort[1:6, ]
head(coef(lmer_mod_slope0)$cohort)

dev.off()

# plot lmers


cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5
cc_half_life(lmer_mod_slope)
cc_half_life(lmer_mod_slope_int)


ggplot(data = augment(lmer_mod_slope0) %>% mutate(year_abn = as.numeric(levels(cohort))[cohort],
                                                       age = time + 5),
       mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point(data = dat) +
  geom_line(aes(y = exp(.fitted)), size = 1) + 
  geom_line(aes(y = exp(.fixed)), size = 1.5, color = "red") + 
  geom_vline(xintercept = cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5) + 
  
  # add fitted values, exponentiated
  geom_line(data = augment(lm_mod0, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = "blue", size = 1.5, inherit.aes = FALSE) + 
  
  scale_color_distiller(palette = "Greens") + 
  
  geom_line(data = augment(lmer_mod_slope) %>% mutate(year_abn = as.numeric(levels(cohort))[cohort]),
            aes(x = age, y = exp(.fixed), group = ), size = 1, color = "pink") +
  
  geom_hline(yintercept = 0.5)


# 




ggplot(data = data.frame(ranef(lmer_mod_slope0)$cohort)) + geom_histogram(aes(x = time))












# old crud


lmer_mod1 # (1 + age|year_abn)
lmer_mod # (0 + age|year_abn)

?isSingular
isSingular(lmer_mod) # FALSE
isSingular(lmer_mod1) # TRUE
summary(lmer_mod)
lmer_mod_coef <- tidy(lmer_mod)
lmer_mod_confint <- confint(lmer_mod) %>%
  as_tibble(rownames = "term_conf") %>%
  rename(lower = "2.5 %",
         upper = "97.5 %")

lmer_mod_coef %>% arrange(desc(effect))
lmer_mod_confint


lmer_mod_coef <- bind_cols(lmer_mod_coef %>% arrange(desc(effect)), lmer_mod_confint)


lmer_mod1_coef <- tidy(lmer_mod1)
lmer_mod1_confint <- confint(lmer_mod1) %>%
  as_tibble(rownames = "term_conf") %>%
  rename(lower = "2.5 %",
         upper = "97.5 %")

lmer_mod1_coef <- bind_cols(lmer_mod1_coef %>% arrange(desc(effect)),
                            lmer_mod1_confint)




summary(lmer_mod1)
coef(lmer_mod)
fixef(lmer_mod)
broom.mixed::tidy(lmer_mod)
test <- tidy(lmer_mod) %>% 
  mutate(lower = estimate - 1.96*std.error,
         upper = estimate + 1.96*std.error)

test[which(test$term == "age"), "estimate"]
test[which(test$term == "age"), "lower"] %>% unlist()
test[which(test$term == "age"), "upper"] %>% unlist()


confint(lmer_mod1)
tidy(lmer_mod1)


lmer_mod_coef
lmer_mod_coef[which(lmer_mod_coef$term == "(Intercept)"), "estimate"]
filter(lmer_mod_coef, term == "(Intercept)")$estimate

fixef(lmer_mod)
(log(0.5) - fixef(lmer_mod)["(Intercept)"])/(fixef(lmer_mod)["age"]) # (log(0.5) - int)/coeff = 30.25
(log(0.5) - fixef(lmer_mod1)["(Intercept)"])/(fixef(lmer_mod1)["age"]) # (log(0.5) - int)/coeff = 30.25

(log(0.5) - filter(lmer_mod_coef, term == "(Intercept)")$estimate) / 
  (filter(lmer_mod_coef, term == "age")$estimate) # (log(0.5) - int)/coeff = 30.25
(log(0.5) - filter(lmer_mod1_coef, term == "(Intercept)")$estimate) / 
  (filter(lmer_mod1_coef, term == "age")$estimate) # (log(0.5) - int)/coeff = 30.25



predict(lmer_mod)


plot(lmer_mod)
plot(lmer_mod1)
initial_area_abn

```

```{r plot-lmers}

show_col(hue_pal()(5))
gg_colors <- hue_pal()(5)
gg_colors[2]

gg_lmer_comparison <- 
  ggplot(data = mutate(augment(lmer_mod_slope0), 
                       year_abn = as.numeric(levels(cohort))[cohort],
                       age = time + 5),
       mapping = aes(x = age, y = proportion, 
                     group = year_abn, color = year_abn)) + 
  theme_classic() + 
  labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land: Shaanxi Province",
       subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + 
  theme(legend.position = "bottom",
        legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  scale_x_continuous(n.breaks = 10) +
  
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", size = 0.75) + 
  geom_point(data = dat) + geom_line(data = dat, size = 1, alpha = 0.25) +
  #geom_line(aes(y = exp(.fitted)), size = 1) + 
  
  
  # lmer_mod_slope0
  geom_line(aes(y = exp(.fixed)), size = 1.5, color = gg_colors[1]) + 
  geom_vline(xintercept = cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5,
             color = gg_colors[1]) + 
  annotate("text", fontface = "bold", angle = 90, size = 3.5,
           x = cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5, 
           y = 0.95, label = paste0(
             "lmer_mod_slope0: \n", 
             round(cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5, 2), 
             " years")) + 

  # lmer_mod_slope
  geom_line(data = augment(lmer_mod_slope) %>% mutate(year_abn = as.numeric(levels(cohort))[cohort]),
            aes(y = exp(.fixed)), size = 1.5, color = gg_colors[2]) + 
  geom_vline(xintercept = cc_half_life(lmer_mod_slope), color = gg_colors[2]) + 
  annotate("text", fontface = "italic", angle = 90, size = 3.5,
           x = cc_half_life(lmer_mod_slope), 
           y = 0.85, label = paste0(
             "lmer_mod_slope: \n", 
             round(cc_half_life(lmer_mod_slope), 2), 
             " years")) + 
  
  # lmer_mod_slope_int
  geom_line(data = augment(lmer_mod_slope_int) %>% mutate(year_abn = as.numeric(levels(cohort))[cohort]),
            aes(y = exp(.fixed)), size = 1.5, color = gg_colors[3]) + 
  geom_vline(xintercept = cc_half_life(lmer_mod_slope_int), color = gg_colors[3]) + 
  annotate("text", fontface = "italic", angle = 90, size = 3.5,
           x = cc_half_life(lmer_mod_slope_int), 
           y = 0.85, label = paste0(
             "lmer_mod_slope_int: \n", 
             round(cc_half_life(lmer_mod_slope_int), 2), 
             " years")) + 
  
  # lm_mod0
  geom_line(data = augment(lm_mod0, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = gg_colors[4], size = 1.5, inherit.aes = FALSE) + 
  geom_vline(xintercept = cc_half_life(lm_mod0, fixed_effect = "time") + 5, color = gg_colors[4]) + 
  annotate("text", fontface = "italic", angle = 90, size = 3.5,
           x = cc_half_life(lm_mod0, fixed_effect = "time") + 5, 
           y = 0.85, label = paste0(
             "lm_mod0: \n", 
             round(cc_half_life(lm_mod0, fixed_effect = "time") + 5, 2), 
             " years")) + 
  
  
  # log log
  geom_line(data = augment(lm_mod_log2, newdata = data.frame(age = 5:30)),
            aes(x = age, y = exp(.fitted)),
            color = gg_colors[5], size = 1.5, inherit.aes = FALSE) + 
  geom_vline(xintercept = (0.5/exp(coef(lm_mod_log2)[1]))^(1/coef(lm_mod_log2)[2]), 
             color = gg_colors[5]) + 
  annotate("text", fontface = "italic", angle = 90, size = 3.5,
           x = (0.5/exp(coef(lm_mod_log2)[1]))^(1/coef(lm_mod_log2)[2]), 
           y = 0.75, label = paste0(
             "lm_mod_log2: \n", 
             round((0.5/exp(coef(lm_mod_log2)[1]))^(1/coef(lm_mod_log2)[2]), 2), 
             " years"))
  # geom_function(fun = function(x) { x^(coef(lm_mod_log2)[2])*exp(coef(lm_mod_log2)[1])},
  #               color = gg_colors[5], size = 1, linetype = "dashed", inherit.aes = FALSE)


gg_lmer_comparison

gg_lmer_comparison + geom_function(fun = function(x) {exp( mean(coef(lm_mod_cohort_slope0), na.rm = TRUE)* (x - 5))},
                color = "blue", size = 1, linetype = "solid", inherit.aes = FALSE)

# save
png(filename = paste0(p_output, "plots/lmer_comparison_s.png"), 
    width = 8, height = 6, units = "in", res = 400)
print(gg_lmer_comparison)
dev.off()


gg_lmer_slope0


gg_lmer_slope0 <- ggplot(data = mutate(augment(lmer_mod_slope0), 
                       year_abn = as.numeric(levels(cohort))[cohort],
                       age = time + 5),
       mapping = aes(x = age, y = proportion, 
                     group = year_abn, color = year_abn)) + 
  theme_classic() + 
  labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land: Shaanxi Province",
       subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + 
  theme(legend.position = "bottom",
        legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  scale_x_continuous(n.breaks = 10) +
  
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "black", size = 0.75) + 
  geom_line(aes(y = exp(.fixed)), size = 1.5, color = gg_colors[1]) + 

  
  geom_point(data = dat) + geom_line(data = dat, size = 1, alpha = 0.25) +
  geom_line(aes(y = exp(.fitted)), size = 1) + 
  
  
  # lmer_mod_slope0
  geom_vline(xintercept = cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5,
             color = gg_colors[1]) + 
  annotate("text", fontface = "italic", angle = 90, size = 3.5,
           x = cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5, 
           y = 0.85, label = paste0(
             "lmer_mod_slope0: \n", 
             round(cc_half_life(lmer_mod_slope0, fixed_effect = "time") + 5, 2), 
             " years"))



# save
png(filename = paste0(p_output, "plots/lmer_slope0_s.png"), 
    width = 7, height = 6, units = "in", res = 400)
print(gg_lmer_slope0)
dev.off()

gg_lmer_comparison
gg_lmer_slope0
```

```{r lmer-area}
# --------------------------------------------------------- #
# start with basic lm
# log(proportion) ~ age + initial_area_abn
# --------------------------------------------------------- #
lm_area_p <- lm(log(proportion) ~ age + initial_area_abn, data = dat)

summary(lm_area_p)
plot(lm_area)
coef(lm_area)
tidy(lm_area)
augment(lm_area)
summary(dat$initial_area_abn)
exp(0.5063)


plot_base <- ggplot(data = persistence_list_b1_s$na_first) + 
  theme_classic() + 
  geom_line(mapping = aes(x = age, y = proportion, 
                          group = year_abn, color = year_abn), 
            size = 1) + 
 labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land: Shaanxi Province",
       subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + theme(legend.position = "bottom") +
  theme(legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  geom_hline(yintercept = 0.5, linetype="dashed", 
             color = "black", size = 0.75)

# initial_area_abn
ggplot(data = dat) + 
  theme_classic() + 
  geom_line(mapping = aes(x = age, y = proportion, 
                          group = year_abn, 
                          color = year_abn), 
            size = 1) + 
 labs(y = "Proportion remaining abandoned", x = "Years since initial abandonment") + 
  scale_color_distiller(palette = "Greens") + 
  geom_point(data = dat %>% mutate(prediction = exp(predict(lm_area_p))),
             mapping = aes(x = age, y = prediction), alpha = 0.5)


# trying cut to make bins
ggplot(data = dat,
       mapping = aes(x = age, y = proportion, 
                     group = year_abn, color = cut(year_abn, breaks = 5))) + 
  theme_classic() + 
  geom_line() + 
  labs(y = "Proportion remaining abandoned", x = "Years since initial abandonment")

ggplot(data = dat %>% mutate(year_abn =
                               as.numeric(levels(year_abn))[year_abn]),
       mapping = aes(x = age, y = proportion, 
                     group = year_abn, color = year_abn_bins)) + 
  theme_classic() + 
  geom_line() + 
  labs(y = "Proportion remaining abandoned", x = "Years since initial abandonment")

#plot relationship between initial area abandoned and the year abandoned
ggplot(data = dat %>% 
         filter(proportion == 1) %>% 
         mutate(year_abn = as.numeric(levels(year_abn))[year_abn]), 
       mapping = aes(x = year_abn, y = initial_area_abn)) + 
  theme_classic() +
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x, mapping = aes(x = year_abn, y = initial_area_abn))



# --------------------------------------------------------- #
# now, 
# log(proportion) ~ age + initial_area_abn
# --------------------------------------------------------- #

lm_area <- lm(log(area_ha) ~ age, data = dat)
summary(lm_area)
plot(lm_area)

ggplot(data = dat %>% mutate(year_abn =
                               as.numeric(levels(year_abn))[year_abn])) + 
  theme_classic() + 
  geom_line(mapping = aes(x = age, y = proportion, 
                          group = year_abn, 
                          color = year_abn), 
            size = 1) + 
 labs(y = "Proportion remaining abandoned", x = "Years since initial abandonment") + 
  scale_color_distiller(palette = "Greens") + 
  geom_point(data = dat %>% mutate(prediction = exp(predict(lm_area_p))),
             mapping = aes(x = age, y = prediction), alpha = 0.5)


hist(dat$proportion)
ggplot(data = dat, mapping = aes(x = log(proportion))) + 
  geom_histogram(binwidth = 0.015)
  






# add to basic lmer, proportions
str(dat)
lmer_mod2 <- lmer(log(proportion) ~ 
       age + initial_area_abn + # fixed effect (i.e. what we're curious about)
       (1 + age|year_abn), # random effects (random intercept, and random slopes)
     data = dat)
summary(lmer_mod2)






lmer_mod_area <- lmer(log(area_ha) ~ 
       age + # fixed effect (i.e. what we're curious about). Do I include a term for initial_area_abn?
       (1 + age|year_abn), # random effects (random intercept, and random slopes)
       data = dat)

# the initial_area_abn should be correlated with each cohort (year_abn), so I'm not sure I need to include it. Better ask Alex about this.

lmer(log(area_ha) ~ age + (1 + initial_area_abn|year_abn), data = persistence_list_b1_s$na_first)
# pull out the initial area abn...^^^

summary(lmer_mod_area)
tidy(lmer_mod_area)


# plot results --------------------------------------------------- #

plot_base <- ggplot(data = persistence_list_b1_s$na_first) + 
  theme_classic() + 
  geom_line(mapping = aes(x = age, y = proportion, 
                          group = year_abn, color = year_abn), 
            size = 1) + 
 labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land: Shaanxi Province",
       subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + theme(legend.position = "bottom") +
  theme(legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  geom_hline(yintercept = 0.5, linetype="dashed", 
             color = "black", size = 0.75)

plot_base +
  
  # lmer_mod
  stat_function(fun = function(x) {exp(filter(lmer_mod_coef, term == "age")$estimate*x + 
                                        filter(lmer_mod_coef, term == "(Intercept)")$estimate)},
                color = gg_colors[2], size = 1, linetype = "solid") + 
  geom_ribbon(data = mutate(data.frame(x = 5:35),
                            lower = exp(filter(lmer_mod_coef, term == "age")$lower*x +
                                          filter(lmer_mod_coef, term == "(Intercept)")$lower),
                            upper = exp(filter(lmer_mod_coef, term == "age")$upper*x +
                                          filter(lmer_mod_coef, term == "(Intercept)")$upper)),
              mapping = aes(ymin = lower, ymax = upper, x = x),
              fill = gg_colors[2], alpha = 0.1, color = gg_colors[2], linetype = "dashed") +
  annotate("text", fontface = "italic", angle = 90,
           x = ((log(0.5) - filter(lmer_mod_coef, term == "(Intercept)")$estimate) /     
                      (filter(lmer_mod_coef, term == "age")$estimate)), 
           y = 0.85, label = "lmer (random slopes): \n34.35 years", size = 3.5) + 
  geom_vline(xintercept = 
               ((log(0.5) - filter(lmer_mod_coef, term == "(Intercept)")$estimate) /     
                  (filter(lmer_mod_coef, term == "age")$estimate)), 
             color = gg_colors[2], linetype = "solid")
    


```

# Check Assumptions

```{r mod_l}
# models
mod_l <- list("lm_mod" = lm_mod, 
              "lm_mod0" = lm_mod0, 
              "lm_mod_log2" = lm_mod_log2,
              "lm_mod_log2_0" = lm_mod_log2_0,
              "lm_mod_cohort_int"  = lm_mod_cohort_int,
              "lm_mod_cohort_slope"  = lm_mod_cohort_slope,
              "lm_mod_cohort_slope0" = lm_mod_cohort_slope0,
              "lm_mod_cohort_x" = lm_mod_cohort_x, 
              "lm_mod_log2_cohort" = lm_mod_log2_cohort,
              "lm_mod_log2_year_abn" = lm_mod_log2_year_abn,
              "lmer_mod_slope" = lmer_mod_slope,
              "lmer_mod_slope_int" = lmer_mod_slope_int,
              "lmer_mod_int" = lmer_mod_int,
              "lmer_mod_slope0" = lmer_mod_slope0
              )



```


```{r AIC}
pairs(dat)
mod_l[[1]]

summary(mod_l[[1]])
glance(mod_l[[1]])

names(mod_l)

lapply(mod_l, formula)
lapply(mod_l, print)

data(lm_mod)
lm_mod$call
lmer_mod$call
lmer_mod %>% formula()
tidy(lmer_mod)
glance(lmer_mod)
augment(lmer_mod)

lapply(mod_l, glance)
summary(lmer_mod_slope0)

install.packages("MuMIn")
library(MuMIn)
MuMIn::r.squaredGLMM(lmer_mod)

plot(lm_mod_log2_cohort)





# calculate AIC and compare across models:

lapply(mod_l, FUN = function(x) {
  glance(x) %>% select(AIC)
}) %>% bind_rows(.id = "mod") %>% arrange(AIC)

# best model
# 1.
lm_mod_log2_cohort

# 2.
lm_mod_cohort_x
```



```{r Oscar}
# chat with Oscar

# best models so far:
lm_mod_log2_cohort
lm_mod_cohort_slope0
lmer_mod_slope0


# plots
lm_mod_cohort_slope0
ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 

  labs(title = "log(proportion) ~ 0 + time:cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + time:cohort, data = dat)), 
                          year_abn = as.numeric(levels(cohort))[cohort], age = time + 5),
            aes(x = age, y = exp(.fitted)))


# lmer_mod_slope0
lmer_mod_slope0
ggplot(data = augment(lmer_mod_slope0) %>% mutate(year_abn = as.numeric(levels(cohort))[cohort],
                                                       age = time + 5),
       mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point(data = dat) +
  geom_line(aes(y = exp(.fitted)), size = 1) + 
  
  # average of the model
  geom_line(aes(y = exp(.fixed)), size = 1.5, color = "red") + 
  
  # lm_mod0
  geom_line(data = augment(lm_mod0, newdata = mutate(data.frame(time = 0:25), age = time + 5)),
            aes(x = age, y = exp(.fitted)),
            color = "blue", size = 1.5, inherit.aes = FALSE) + 
  
  scale_color_distiller(palette = "Greens") + 
  
  geom_hline(yintercept = 0.5)


# log log 0
lm_mod_log2_cohort
ggplot(data = dat, mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  labs(title = "log(proportion) ~ 0 + log(time_1):cohort") + 
  geom_line(data = mutate(augment(lm(log(proportion) ~ 0 + log(time_1):cohort, 
                                     data = mutate(dat, time_1 = time + 1))), 
                          year_abn = as.numeric(levels(cohort))[cohort]),
            aes(x = exp(`log(time_1)`) - 1 + 5, y = exp(.fitted))) + 
  scale_x_continuous(breaks = seq(from = 0, to = 30, by = 5))









# diagnostic plots:
par(mfrow = c(2,2))
plot(lm_mod_cohort_slope0)
plot(lm_mod_log2_cohort)

dev.off()

plot(lmer_mod_slope0)
qqnorm(residuals(lmer_mod_slope0))

dat[17:19, ]






# residual plots




```

```{r check-assumptions}

# ------- check assumptions of the lm ------------------------------------------------- #
lm_mod
par(mfrow = c(2,2))
plot(mod_l[[9]])
residuals(lm_mod)
broom::augment(lm_mod)

# plot model diagnostics:
par(mfrow = c(2,2)) # create a plotting panel with two rows and two columns (allowing you to put 4 plots on the same plotting panel)
plot(lm_mod) # check for violations of model assumptions

dev.off()

# description of the four diagnostic plots from: http://www.sthda.com/english/articles/39-regression-model-diagnostics/161-linear-regression-assumptions-and-diagnostics-in-r-essentials/
# 1. Residuals vs Fitted. Used to check the linear relationship assumptions. A horizontal line, without distinct patterns is an indication for a linear relationship, which is good.
# 2. Normal Q-Q. Used to examine whether the residuals are normally distributed. Good if residuals points follow the straight dashed line.
# 3. Scale-Location (or Spread-Location). Used to check the homogeneity of variance of the residuals (homoscedasticity). Horizontal line with equally spread points is a good indication of homoscedasticity.
# 4. Residuals vs Leverage. Used to identify influential cases, that is extreme values that might influence the regression results when included or excluded from the analysis. 



# ------------------------------------- Umesh's notes on assumptions ------------------------------------------------ #
# assumptions to check:
# a. LINEAR RELATIONSHIP between the predictor (x) and the response (y); i.e., y = mx + C
# b. the data points are STATISTICALLY INDEPENDENT (study design issue)
# c. no outliers influence the relationship unduly (look at the Residuals vs Leverage plot; also the Normal Q-Q plot)
# d. CONSTANT, or EQUAL VARIANCE (fancily called HOMOSCEDASTICITY; look at the Residuals vs Fitted plot)
# e. ERRORS ARE NORMALLY DISTRIBUTED (look at the Normal Q-Q plot)

# ------------------ a. Linear relationship plot to check.
ggplot(data = dat, mapping = aes(x = age, y = log(proportion), group = year_abn, color = year_abn)) + 
  theme_classic() + 
  geom_point(size = 1.25) + 
  labs(title = "Persistence of Abandoned Land") + 
  scale_color_distiller(palette = "Greens") + 
  theme(legend.position = "bottom", legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  geom_smooth(method = "lm", aes(x = age, y = log(proportion)), se = FALSE, inherit.aes = FALSE, color = "blue") + 
  geom_hline(yintercept = log(0.5), linetype = "dashed", color = "red", size = 0.75)
# yes?

# ------------------ b. Statistically independent? I don't think the points are independent. Each cohort of abandoned land (i.e. land that was all initially abandoned in a give year) can constitute a group. This is especially the case if I'm looking at area, rather than proportion.
pairs(dat)
plot(proportion ~ year_abn, data = dat)
plot(area_ha ~ initial_area_abn, data = dat)
plot(proportion ~ initial_area_abn, data = dat)

# If the points are not independent, and they belong to a particular cohort of abandoned land, I have two options: 
# add cohorts as a categorical fixed effect, calculating either unique intercepts, unique slopes, or both depending on the identity of the cohort. 
# Or, I could model these cohort groups as random effects (meaning that we expect the behavior or each cohort group to be drawn from a normal distribution with a particular mean and standard deviation), and move to a linear mixed effects model next. 

# other option: Bayesian hierarchical model. 
install.packages("brms") # Hamiltonian monte carlo.


# continuing with Umesh's diagnostics:
# ------------------ c.
# c. no outliers influence the relationship unduly (look at the Residuals vs Leverage plot; also the Normal Q-Q plot)
# the residuals should lie along the dotted line in the QQ plot, and the leverage plot should look straight as well.


# ------------------ d.
# d. CONSTANT, or EQUAL VARIANCE (fancily called HOMOSCEDASTICITY; look at the Residuals vs Fitted plot)
# this first plot should have a pretty horizontal line across it, with good horizontal spread, and vertical spread. An even distribution along the x axis (fitted values) across values of y indicates a lack of *bias*. An even spread across the y axis across values of x shows homoscedasticity. Constant variance among residuals.  

# ------------------ e.
# e. Error ARE NORMALLY DISTRIBUTED (look at the Normal Q-Q plot)
# looks good

plot(lm_mod, 2)

# can also 

# check normality of residuals
hist(residuals(lm_mod)) # or
ggplot(data = data.frame(.resid = residuals(lm_mod)), aes(x = .resid)) + geom_histogram() # looks normal to me?


ggplot(data = augment(lmer_mod_offset), aes(x = .resid)) + geom_histogram() # looks normal to me?
ggplot(data = augment(lmer_mod_offset), aes(x = `log(proportion)`)) + geom_histogram() # looks normal to me?


plot(lmer_mod)
qqnorm(residuals(lmer_mod))

plot(lmer_mod_offset)
qqnorm(residuals(lmer_mod_offset))



# --------------------------------- checking assumptions for lmers -------------------------- #
names(mod_l)
plot(mod_l$lm_mod_cohort_slope0)
plot(mod_l$lm_mod_log2_cohort)

qqnorm(residuals(mod_l[[8]]))
tidy(mod_l[[8]])
coef(mod_l[[8]])
augment(mod_l[[8]])
plot(mod_l$lm_mod_cohort_slope0$residuals)

ggplot(data = augment(mod_l$lm_mod_cohort_slope0), aes(x = .std.resid)) + geom_histogram() # looks normal to me?
ggplot(data = augment(mod_l[[8]]), aes(x = `log(proportion)`)) + geom_histogram() # this doesn't look all that normal to me

# leverage
ggplot(data = augment(mod_l[[8]]), aes(x = .hat)) + geom_histogram()



lapply(mod_l, summary)
```

```{r car}
# resource from Oscar: 
# https://dss.princeton.edu/training/Regression101R.pdf

install.packages("car")
library(car)

# plot model diagnostics:
par(mfrow = c(2,2)) # create a plotting panel with two rows and two columns (allowing you to put 4 plots on the same plotting panel)
plot(lm_log2_l[[9]]) # check for violations of model assumptions


# Now, with car
residualPlots(lm_log2_l[[9]])


# remove outliers, rerun the model

lm_trim <- lm(log(proportion) ~ 0 + log(time_1):cohort, 
              data = filter(dat, cohort %in% c(1988:2010)))
plot(lm_trim)
plot(lm_trim)


ggplot(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point(data = dat) + scale_color_distiller(palette = "Greens") + 
  geom_line(data = mutate(augment(lm_trim, newdata = dat0 %>% filter(cohort %in% c(1988:2010))),
                          proportion = exp(.fitted)))
```


# Coefficient plots
```{r coef-over-time-s}
tidy(lm_mod_log2_cohort) %>%
  mutate(year_abn = 1988:2013)

tidy(lm_mod_log2_cohort, conf.int = TRUE) %>% mutate(year_abn = 1988:2013) %>% print(n = 30)

gg_coef_time_s <- ggplot(data = tidy(lm_mod_log2_cohort, conf.int = TRUE) %>% mutate(year_abn = 1988:2013),
       mapping = aes(x = year_abn, y = estimate)) + 
  theme_classic() +
  geom_point() + 
  geom_errorbar(mapping = aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  labs(y = "Model Coefficient (Power Term)",
       x = "Cohort (Year of Initial Abandonment)", 
       title = paste0("Model Coefficients: ", site_df$description[indx]),
       subtitle = "log(proportion) ~ 0 + log(time_1):cohort",
       caption = "Error bars represent 95% confidence interval"
       # fill = "Site"
       ) + 
  theme(#axis.text.x = element_text(angle = 320, vjust = 1, hjust = 0), 
        legend.position = "none", plot.caption = element_text(face = "italic"),
        plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm")) + 
  geom_smooth(method = "lm")


# half lives over time
# y = x ^ coef
# 0.5 = x ^ coef
tidy(lm_mod_log2_cohort, conf.int = TRUE) %>% 
  mutate(hl = 0.5^(1/estimate),
         hl_low = 0.5^(1/conf.low),
         hl_high = 0.5^(1/conf.high),
         p_25 = 20^(estimate)
         )

half_life_df <- mutate(tidy(lm_mod_log2_cohort, conf.int = TRUE), 
                       hl = 0.5^(1/estimate),
                       hl_low = 0.5^(1/conf.low),
                       hl_high = 0.5^(1/conf.high),
                       year_abn = 1988:2013,
                       p_25 = 21^(estimate),
                       p_25_low = 21^(conf.low),
                       p_25_high = 21^(conf.high)
                       )

half_life_df %>% print(n = 30)


gg_prop_at_25_s <- 
  ggplot(data = half_life_df,
       mapping = aes(x = year_abn, y = p_25)) + 
  theme_classic() +
  geom_point() + 
  geom_errorbar(mapping = aes(ymin = p_25_low, ymax = p_25_high), width = 0.3) +
  labs(y = "Proportion remaining after 25 years", x = "Cohort (Year of Initial Abandonment)", 
       title = paste0("Decay rates over time: ", site_df$description[indx]),
       subtitle = "log(proportion) ~ 0 + log(time_1):cohort",
       caption = "Error bars represent 95% confidence interval") + 
  geom_smooth(method = "lm") +
  # scale_y_continuous(limits = c(0, 250)) +
  theme(plot.caption = element_text(face = "italic"),
        plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm"))


gg_half_life_s <- 
  ggplot(data = half_life_df,
       mapping = aes(x = year_abn, y = hl)) + 
  geom_point() + 
  geom_errorbar(mapping = aes(ymin = hl_low, ymax = hl_high), width = 0.3) +
  labs(y = "Half-Life (Years)", x = "Cohort (Year of Initial Abandonment)", 
       title = paste0("Model Coefficients: ", site_df$description[indx]),
       subtitle = "log(proportion) ~ 0 + log(time_1):cohort",
       caption = "Error bars represent 95% confidence interval") + 
  theme( plot.caption = element_text(face = "italic"),
        plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm")) + 
  #geom_smooth(method = "lm") + 
  scale_y_continuous(limits = c(0, 250))



# save
png(filename = paste0(p_output, "plots/log2_cohort_coef_time_s.png"), 
    width = 6, height = 5, units = "in", res = 400)
print(gg_coef_time_s)
dev.off()

# save
png(filename = paste0(p_output, "plots/log2_cohort_prop_at_25_s.png"), 
    width = 6, height = 5, units = "in", res = 400)
print(gg_prop_at_25_s)
dev.off()


```


```{r coef-all-sites}
# extract coefficients for log-log-cohort model for all sites
coef_df <- lapply(lm_log2_l, FUN = function(x) {
  tidy(x, conf.int = TRUE) %>% mutate(year_abn = 1988:2013)
}) %>% bind_rows(.id = "site")

# plot all coefficients
ggplot(data = coef_df %>% filter(),
       mapping = aes(x = year_abn, y = estimate, group = site, color = site)) + 
  theme_classic() +
  geom_point(position = position_jitter()) + 
  #geom_errorbar(mapping = aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  labs(y = "Model Coefficient (Power Term)",
       x = "Cohort (Year of Initial Abandonment)", 
       title = "Model Coefficients",
       subtitle = "log(proportion) ~ 0 + log(time_1):cohort",
       caption = "Error bars represent 95% confidence interval",
       color = "Site") + 
  theme(#axis.text.x = element_text(angle = 320, vjust = 1, hjust = 0), 
        legend.position = "bottom",
        plot.caption = element_text(face = "italic"),
        plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm")) + 
  geom_smooth(method = "lm", se = FALSE) # + 
  # plot line manually
  # geom_abline(slope = 8.812e-04 - 5.977038e-03, intercept = 1.004106e+01)


# with quadrants:
gg_log2_cohort_coef_all <- ggplot(data = coef_df %>% 
       mutate(site_n = rep(1:11, each = 26), 
              cluster = cut(site_n, breaks = 4),
              # to customize which site is in which quadrant
              cluster1 = ifelse(site_n %in% c(1, 5, 10), "a",
                               ifelse(site_n %in% c(2, 6, 11), "b",
                                      ifelse(site_n %in% c(3, 7, 9), "c", 
                                             ifelse(site_n %in% c(4, 8), "d", NA))))),
       mapping = aes(x = year_abn, y = estimate, group = site, color = site)) + 
  theme_classic() +
  geom_point(position = position_jitter()) + 
  #geom_errorbar(mapping = aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  labs(y = "Model Coefficient (Power Term)",
       x = "Cohort (Year of Initial Abandonment)", 
       title = "Model Coefficients",
       subtitle = "log(proportion) ~ 0 + log(time_1):cohort",
       #caption = "Error bars represent 95% confidence interval",
       color = "Site") + 
  theme(#axis.text.x = element_text(angle = 320, vjust = 1, hjust = 0), 
        legend.position = "bottom",
        plot.caption = element_text(face = "italic"),
        plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm")) + 
  geom_smooth(method = "lm", se = FALSE)  + 
  facet_wrap(vars(cluster)) + 
  theme(strip.background = element_blank(), strip.text = element_blank())

# save:
png(filename = paste0(p_output, "plots/log2_cohort_coefs_all.png"), 
    width = 7, height = 7, units = "in", res = 400)
print(gg_log2_cohort_coef_all)
dev.off()
```


```{r mean-coef-per-site}
# mean_coef per site
mean_coef_df <- coef_df %>% 
  group_by(site) %>% 
  summarise(mean = mean(estimate, na.rm = TRUE),
            mean_low = mean(conf.low, na.rm = TRUE),
            mean_high = mean(conf.high, na.rm = TRUE)) %>% 
  mutate(term = "coef")

mean_coef_df <- bind_rows(mean_coef_df, 
                          lapply(c(21, 46), FUN = function(x) {
                            mean_coef_df %>%
                              mutate(mean = x^(mean),
                                     mean_low = x^(mean_low),
                                     mean_high = x^(mean_high),
                                     term = paste0("p", x + 4))
                            }))

# quasi half-lives
lapply(lm_log2_l, FUN = function(i) cc_half_life(i, confint = TRUE)) %>% bind_rows(.id = "site")


gg_mean_coef_site <- ggplot(data = mean_coef_df %>% filter(term == "coef"),
                            mapping = aes(x = mean, y = site)) + 
  theme_classic() +
  geom_point() + 
  geom_errorbar(mapping = aes(xmin = mean_low, xmax = mean_high), width = 0.3) +
  labs(x = "Mean Coefficient (Power Term)", y = NULL, 
       title = "Mean Decay Rates By Site",
       subtitle = "log(proportion) ~ 0 + log(time_1):cohort",
       caption = "Error bars represent the mean of high and low cofficients from 95% confidence interval") + 
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme(plot.caption = element_text(face = "italic"))

# save
png(filename = paste0(p_output, "plots/log2_cohort_coefs_by_site.png"), 
    width = 5, height = 6, units = "in", res = 400)
print(gg_mean_coef_site)
dev.off()

gg_coef_p25_by_site <- ggplot(data = mean_coef_df,
       mapping = aes(y = mean, x = site, col = site)) + 
  theme_classic() +
  geom_point() + 
  geom_errorbar(mapping = aes(ymin = mean_low, ymax = mean_high), width = 0.3) +
  labs(y = NULL, x = "Site", 
       title = "Decay rates by site",
       subtitle = "log(proportion) ~ 0 + log(time_1):cohort",
       caption = "Error bars represent the mean of 95% confidence interval coefficient estimates") + 
  # geom_hline(yintercept = 0, linetype = "dashed") +
  facet_grid(rows = vars(term), scales = "free", #switch = "y",
             labeller = labeller(term = c(coef = "Mean Coefficient", #old = "new",
                                          p25 = "Proportion remaining \nafter 25 years",
                                          p50 = "Proportion remaining \nafter 50 years"))) + 
  theme(plot.caption = element_text(face = "italic"),
        axis.text.x = element_text(angle = 320, vjust = 1, hjust = 0),
        legend.position = "none"#, plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm")
        )



# save
png(filename = paste0(p_output, "plots/log2_cohort_coef_p25_by_site.png"), 
    width = 6, height = 7, units = "in", res = 400)
print(gg_coef_p25_by_site)
dev.off()
```


```{r coef-rate-of-change-per-site}
lm_coef <- lm(estimate ~ year_abn * site - 1, data = coef_df)

coef(lm_coef)[1]
as.numeric(tidy(lm_coef, conf.int = TRUE)[1, "estimate"])

coef_df2 <- tidy(lm_coef, conf.int = TRUE) %>%
  mutate(termm = ifelse(grepl("year_abn", term), "slope", "int"),
         site = gsub("site|year_abn:", "", term)) %>% 
  pivot_wider(id_cols = c(site, termm), names_from = termm, values_from = c(estimate, conf.low, conf.high)) %>% 
  mutate(
    estimate_slope = ifelse(site == "belarus", 0, estimate_slope),
    conf.low_slope = ifelse(site == "belarus", 0, conf.low_slope),
    conf.high_slope = ifelse(site == "belarus", 0, conf.high_slope),
    estimate_slope = estimate_slope + as.numeric(tidy(lm_coef, conf.int = TRUE)[1, "estimate"]),
    conf.low_slope = conf.low_slope + as.numeric(tidy(lm_coef, conf.int = TRUE)[1, "conf.low"]),
    conf.high_slope = conf.high_slope + as.numeric(tidy(lm_coef, conf.int = TRUE)[1, "conf.high"])
    ) %>% 
  filter(site != "year_abn")


#

gg_coef_rate_change_by_site <- ggplot(data = coef_df2, 
                                      aes(x = site, y = estimate_slope, color = site, fill = site)) + 
  theme_classic() +
  geom_col() + 
  geom_errorbar(mapping = aes(ymin = conf.low_slope, ymax = conf.high_slope), width = 0.3) +
  labs(title = "Changes in Decay Rate over time, by site", 
       x = "Site", y = "Rate of Change of Model Coefficients (per year)") + 
  theme(axis.text.x = element_text(angle = 320, vjust = 1, hjust = 0), legend.position = "none",
        plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm"))

png(filename = paste0(p_output, "plots/log2_cohort_coefs_rate_of_change_by_site.png"), 
    width = 7, height = 6, units = "in", res = 400)
print(gg_coef_rate_change_by_site)
dev.off()

```


# old half-life stuff
```{r half-life-over-time-by-site}
# build data frames of half lives for each model


grep("time", lapply(mod_l, formula))
grep("lmer", names(mod_l))

mod_l[6:9]


hl1 <- lapply(mod_l[6:8], FUN = function(x) {
  coef(x)$cohort %>%
  as_tibble(rownames = "cohort") %>%
  rename(int = "(Intercept)", coeff = age) %>% 
  mutate(half_life = (log(0.5) - int)/coeff,
         year_abn = as.numeric(cohort),
         mod = names(x))
})

bind_rows(hl1, .id = "mod")

hl4 <- coef(mod_l[[9]])$cohort %>%
  as_tibble(rownames = "year_abn") %>%
  rename(coeff = time) %>% 
  mutate(half_life = log(0.5)/coeff + 5,
         year_abn = as.numeric(year_abn),
         mod = "lmer_mod_slope0")

lmer_mod1_half_lives <- coef(lmer_mod1)$year_abn %>%
  as_tibble(rownames = "year_abn") %>%
  rename(int = "(Intercept)", coeff = age) %>% 
  mutate(half_life = (log(0.5) - int)/coeff,
         year_abn = as.numeric(year_abn),
         mod = "lmer1")

mod_l_coef <- lapply(mod_l, FUN = function(x) if(class(x) == "lm") {coef(x)} else {coef(x)[[1]]})





mod_l[c(1,3,7)]

# combine the data.frames to compare models
half_life_mod_comparison <- bind_rows(lmer_mod_half_lives, 
                                      lmer_mod_slope0_half_lives,
                                      lmer_mod1_half_lives, 
                                      select(mod_ind_df, !cohort)) %>% 
  mutate(mod = as_factor(mod))



gg_mod_comp_s <- 
  ggplot(data = half_life_mod_comparison) + theme_classic() +
  geom_smooth(method = "lm", formula = y ~ x,
              mapping = aes(x = year_abn, y = half_life, 
                            fill = mod, color = mod),
              #color = "blue", fill = "gray", 
              size = 1.5, alpha = 0.25, show.legend = FALSE) +
  geom_point(mapping = aes(x = year_abn, y = half_life, color = mod), 
             alpha = 078) + 
  labs(y = "Half life (years)", x = "Year Abandoned", 
       title = "Persistence of Abandoned Land", subtitle = site_df$description[9],
       color = "Model Type")


# save plots

png(filename = paste0(p_output, "plots/half_life_cohorts_s.png"), 
    width = 6, height = 5, units = "in", res = 400)
print(gg_mod_comp_s)
dev.off()
```


```{r plot-half-life-lm-vs-lmer}
# plot half-life for full site:
summary(lmer_mod)
summary(lmer_mod1)

lmer_mod_coef
lmer_mod1_coef
show_col(hue_pal()(3))
gg_colors <- hue_pal()(3)
gg_colors[2]

gg_half_life_lm_s

gg_half_life_lm_lmer_s <- 
  ggplot(data = persistence_list_b1_s$na_first) + 
  # base plot
  theme_classic() + 
  geom_line(mapping = aes(x = age, y = proportion, 
                          group = year_abn, color = year_abn), 
            size = 1, alpha = 0.25) + 
  geom_point(mapping = aes(x = age, y = proportion,
                           group = year_abn, color = year_abn), 
            size = 1) + 
  labs(y = "Proportion remaining abandoned", 
       x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land: Shaanxi Province",
       subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + theme(legend.position = "bottom") +
  theme(legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  geom_hline(yintercept = 0.5, linetype="dashed", 
             color = "black", size = 0.75) + 
  
  # lmer_mod
  stat_function(fun = function(x) {exp(filter(lmer_mod_coef, term == "age")$estimate*x + 
                                        filter(lmer_mod_coef, term == "(Intercept)")$estimate)},
                color = gg_colors[2], size = 1, linetype = "solid") + 
  geom_ribbon(data = mutate(data.frame(x = 5:35),
                            lower = exp(filter(lmer_mod_coef, term == "age")$lower*x +
                                          filter(lmer_mod_coef, term == "(Intercept)")$lower),
                            upper = exp(filter(lmer_mod_coef, term == "age")$upper*x +
                                          filter(lmer_mod_coef, term == "(Intercept)")$upper)),
              mapping = aes(ymin = lower, ymax = upper, x = x),
              fill = gg_colors[2], alpha = 0.1, color = gg_colors[2], linetype = "dashed") +
  annotate("text", fontface = "italic", angle = 90,
           x = ((log(0.5) - filter(lmer_mod_coef, term == "(Intercept)")$estimate) /     
                      (filter(lmer_mod_coef, term == "age")$estimate)), 
           y = 0.85, label = "lmer (random slopes): \n34.35 years", size = 3.5) + 
  geom_vline(xintercept = 
               ((log(0.5) - filter(lmer_mod_coef, term == "(Intercept)")$estimate) /     
                  (filter(lmer_mod_coef, term == "age")$estimate)), 
             color = gg_colors[2], linetype = "solid") + 
    
  # lmer_mod1  
  stat_function(fun = function(x) {exp(filter(lmer_mod1_coef, term == "age")$estimate*x + 
                                        filter(lmer_mod1_coef, term == "(Intercept)")$estimate)},
                color = gg_colors[3], size = 1, linetype = "solid") + 
  geom_ribbon(data = mutate(data.frame(x = 5:35),
                            lower = exp(filter(lmer_mod1_coef, term == "age")$lower*x +
                                          filter(lmer_mod1_coef, term == "(Intercept)")$lower),
                            upper = exp(filter(lmer_mod1_coef, term == "age")$upper*x +
                                          filter(lmer_mod1_coef, term == "(Intercept)")$upper)),
              mapping = aes(ymin = lower, ymax = upper, x = x),
              fill = gg_colors[3], alpha = 0.1, color = gg_colors[3], linetype = "dashed") +
  annotate("text", fontface = "italic", angle = 90,
           x = ((log(0.5) - filter(lmer_mod1_coef, term == "(Intercept)")$estimate) /     
                      (filter(lmer_mod1_coef, term == "age")$estimate)), 
           y = 0.86, label = "lmer (random int. & slopes): \n30.26 years", size = 3.5) + 
  geom_vline(xintercept = 
               ((log(0.5) - filter(lmer_mod1_coef, term == "(Intercept)")$estimate) /     
                  (filter(lmer_mod1_coef, term == "age")$estimate)), 
             color = gg_colors[3], linetype = "solid") + 
  
  # lmer_mod2
  stat_function(fun = function(x) {exp(filter(lmer_mod1_coef, term == "age")$estimate*x + 
                                        filter(lmer_mod1_coef, term == "(Intercept)")$estimate)},
                color = gg_colors[3], size = 1, linetype = "solid") + 
  
  # lm
  stat_function(fun = function(x) {exp(lm_mod$coefficients["age"]*x +
                                         lm_mod$coefficients["(Intercept)"])},
                color = gg_colors[1], size = 1) + 
  annotate(x = ((log(0.5) - lm_mod$coefficients["(Intercept)"]) /     
                      (lm_mod$coefficients["age"])), 
           y = 0.88, label = "lm: \n23.17 years", 
           "text", fontface = "italic", angle = 90, size = 3.5) + 
  geom_vline(xintercept = ((log(0.5) - lm_mod$coefficients["(Intercept)"]) /
                             (lm_mod$coefficients["age"])), 
             color = gg_colors[1]) + 
  geom_ribbon(data = mutate(data.frame(x = 5:35),
                            lower = exp(confint(lm_mod)["age", 1]*x +
                                          confint(lm_mod)["(Intercept)", 1]),
                            upper = exp(confint(lm_mod)["age", 2]*x +
                                          confint(lm_mod)["(Intercept)", 2])),
              mapping = aes(ymin = lower, ymax = upper, x = x),
              fill = gg_colors[1], alpha = 0.1, color = gg_colors[1], linetype = "dashed")

  
gg_half_life_lm_lmer_s

# save
png(filename = paste0(p_output, "plots/half_life_lm_lmer_s.png"), 
    width = 7, height = 6, units = "in", res = 400)
print(gg_half_life_lm_lmer_s)
dev.off()
```


```{r half-life-lm-all-sites}

# ------- calc lm for each site ------------------------------------------------- #
# calculate exponential decay curve for each site:

decay_models <- lapply(1:11, FUN = function(i) {
  load(file = paste0(p_output, "abn_dat_products", blip_label, site_df$label[i], ".rds"), verbose = TRUE)
# eval(parse(text = paste0("persistence_list", blip_label, site_label)))
  
  assign("persistence_list", eval(parse(text = paste0("persistence_list", blip_label, site_df$label[i]))))
  
  model <- lm(log(proportion) ~ age, data = persistence_list$na_first)

})

names(decay_models) <- site_df$site
plot(decay_models$shaanxi)
broom::tidy(decay_models$shaanxi)
broom::tidy(mod_decay_s)


# make a data.frame of coefficients
decay_models_coeff <- lapply(decay_models, coefficients) %>% 
  bind_rows(.id = "site") %>%
  rename(int = "(Intercept)", coeff = age) %>% 
  mutate(half_life = (log(0.5) - int)/coeff)

# calculate confidence intervals
conf <- lapply(decay_models, FUN = function(i) {
  conf <- confint(i) %>% 
    as_tibble(rownames = "rownames") %>% 
    mutate(rownames = ifelse(rownames == "age", "coeff", "int")) %>%
    rename("lower" = "2.5 %", "upper" = "97.5 %") %>% 
    pivot_wider(names_from = rownames, values_from = c(lower, upper))
}) %>% bind_rows(.id = "site")

# join data.frames
decay_models_coeff <- left_join(decay_models_coeff, conf, by = "site") %>%
  mutate(half_life_lower = (log(0.5) - lower_int)/lower_coeff,
         half_life_upper = (log(0.5) - upper_int)/upper_coeff)


# plot
gg_half_life_all_sites <- ggplot(data = decay_models_coeff) + 
  theme_classic() + 
  geom_col(mapping = aes(x = site, y = half_life, fill = site)) + 
  geom_errorbar(mapping = aes(x = site, ymin = half_life_lower, ymax = half_life_upper, #color = site
                              ), width = 0.3) +
  labs(y = "Half-life (years)",
       x = "Site", 
       title = "Half-life of abandoned agricultural land",
       caption = "Error bars represent 95% confidence interval",
       fill = "Site") + 
  theme(axis.text.x = element_text(angle = 320, vjust = 1, hjust = 0), 
        legend.position = "none", plot.caption = element_text(face = "italic"),
        plot.margin = unit(c(0.4, 0.5, 0.2, 0.2), "cm"))




# save plots
# ----------------------------------- save ------------------------------------------------- #

png(filename = paste0(p_output, "plots/half_life_all_sites.png"), 
    width = 5, height = 5, units = "in", res = 400)
print(gg_half_life_all_sites)
dev.off()






# old ----------------------------------------------------------------------- #

# calculating the average proportion at each time_abn, across all sites.
# this replicated df_av, but for all sites

mean_persistence_df <- lapply(1:11, FUN = function(i) {
  load(file = paste0(p_output, "abn_dat_products", blip_label, site_df$label[i], ".rds"), verbose = TRUE)
# eval(parse(text = paste0("persistence_list", blip_label, site_label)))
  
  assign("persistence_list", eval(parse(text = paste0("persistence_list", blip_label, site_df$label[i]))))
  
  df_tmp <- persistence_list$na_last %>%
    mutate(year_abn = as_factor(year_abn)) %>%
    group_by(time_abn) %>%
    summarise(count = n(),
              sd = sd(proportion, na.rm = TRUE), 
              se = sd/sqrt(count),
              mean_proportion = mean(proportion), 
              year_abn = as_factor("mean")) %>%
    mutate(site = site_df$site[i])
  
  df_tmp
})

mean_persistence_df <- bind_rows(mean_persistence_df)

# save
write_csv(mean_persistence_df, path = paste0(p_dat_derived, "mean_persistence_df.csv"))

# re load
mean_persistence_df <- read.csv(file = paste0(p_dat_derived, "mean_persistence_df.csv"))




# plotting the average persistence over time
gg_mean_persistence_all <- ggplot(data = mean_persistence_df) +
  labs(title = "Mean persistence over time", x = "Time abandoned (years)", 
       y = "Mean proportion of abandoned land remaining abandoned") + 
  geom_line(mapping = aes(x = time_abn, y = mean_proportion, group = site, color = site), size = 1) + 
  theme_classic() + theme(legend.position = "right")

# save
png(filename = paste0(p_output, "plots/mean_persistence_all.png"), 
    width = 7, height = 5, units = "in", res = 400)
print(gg_mean_persistence_all)
dev.off()

mean_persistence_df %>% filter(mean_proportion < 0.52 & mean_proportion > 0.49) %>% select(site, time_abn, mean_proportion) %>% as.data.frame()



```


```{r half-life-lm-bins}
# ----------------------------------------------------------------------------------- #
# plot persistence for bins of year abandoned
# bins for year_abn
test$year_abn %>% unique %>% length 
test <- persistence_list$na_first
test <- test %>% 
      mutate(year_abn_bins = ifelse(year_abn >= 1988 & year_abn <= 1993, "1988-1993",
                           ifelse(year_abn >= 1994 & year_abn <= 1998, "1994-1998",
                                  ifelse(year_abn >= 1999 & year_abn <= 2003, "1999-2003",
                                         ifelse(year_abn >= 2004 & year_abn <= 2008, "2004-2008",
                                                ifelse(year_abn >= 2009 & year_abn <= 2013, "2009-2013", NA)
                                         ))))) %>%
      mutate(year_abn_bins = as_factor(year_abn_bins))

year_abn_bins_av <- test %>% group_by(year_abn_bins, age) %>% summarise(mean_proportion = mean(proportion), count = n())


ggplot(data = year_abn_bins_av) + 
  theme_classic() + 
  labs(y = "Proportion remaining abandoned", #x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land", color = "Year Abandoned") + 
  # scale_color_distiller(palette = "Greens") + theme(legend.position = "bottom") +
  theme(legend.text = element_text(angle = 320, vjust = 1, hjust = 0),
        legend.position = "bottom") + 
  geom_line(size = 1.25,
            mapping = aes(x = age, 
                   # x = age, # (replicates the plot from na_last, which is called time_abn)
                   y = mean_proportion, group = year_abn_bins, color = year_abn_bins))


```


```{r half-life-lm-indiv-cohorts}
# half-life for each individual cohort for a particular site
persistence_list$na_first$year_abn %>% unique() %>% length()

# load data set, assign to generic "persistence_list"
load(file = paste0(p_output, "abn_dat_products", blip_label, "_s", ".rds"), verbose = TRUE)
assign("persistence_list", eval(parse(text = paste0("persistence_list", blip_label, "_s"))))


# run an lm() for each cohort, individually
mod_ind <- lapply(1988:2013, FUN = function(i) {
  model <- lm(log(proportion) ~ age, data = filter(persistence_list_b1_s$na_first, year_abn == i))
})

names(mod_ind) <- paste0("y", 1988:2013)

mod_ind_df <- lapply(mod_ind, coefficients) %>% bind_rows(.id = "cohort") %>%
  rename(int = "(Intercept)", coeff = age) %>% 
  mutate(half_life = (log(0.5) - int)/coeff,
         year_abn = gsub("y", "", cohort),
         year_abn = as.numeric(year_abn),
         mod = "lm")

# note: this is the same as the lm_mod_cohort_x 

mod_ind_df
coef(lm_mod_cohort_x)["age"]
coef(lm_mod_cohort_x)[grep("1990", names(coef(lm_mod_cohort_x)))]
tidy(lm_mod_cohort_x) %>% filter(grepl("age", term)) %>% mutate(new = estimate + coef(lm_mod_cohort_x)["age"])


gg_half_life_over_time_s <- ggplot(data = mod_ind_df) + theme_classic() +
  geom_smooth(method = "lm", formula = y ~ x,
              mapping = aes(x = year_abn, y = half_life),
              color = "blue", fill = "gray", size = 1.5, alpha = 0.5) + 
  geom_point(mapping = aes(x = year_abn, y = half_life)) + 
  labs(y = "Half life (years)", x = "Year Abandoned", 
       title = "Persistence of Abandoned Land", subtitle = site_df$description[9],
       color = "Year Abandoned")
  


ggplot(data = persistence_list_b1_s$na_first) + 
  theme_classic() + 
  geom_point(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn), size = 1.25) + 
  labs(y = "Proportion remaining abandoned", x = "Years since initial abandonment", 
       title = "Persistence of Abandoned Land", subtitle = NULL, #subtitle,
       color = "Year Abandoned") + 
  scale_color_distiller(palette = "Greens") + theme(legend.position = "bottom") +
  theme(legend.text = element_text(angle = 320, vjust = 1, hjust = 0)) + 
  geom_hline(yintercept = 0.5, linetype="dashed", color = "red", size = 0.75) + 
  # make and add confidence interval
  geom_ribbon(data = mutate(data.frame(x = 5:30),
                            lower = exp(confint(mod_decay_s)[2, 1]*x + confint(mod_decay_s)[1, 1]),
                            upper = exp(confint(mod_decay_s)[2, 2]*x + confint(mod_decay_s)[1, 2])),
              mapping = aes(ymin = lower, ymax = upper, x = x), 
              fill = "dark gray", alpha = 0.5, color = "dark gray", linetype = "solid") +
  stat_function(fun = function(x) {exp(mod_decay_s$coefficients[2]*x + mod_decay_s$coefficients[1])}, color = "dark blue", size = 1) + 
  annotate("text", x = 24, y = 0.8, label = "half-life = 23.17 years", fontface = "italic", angle = 90) + 
  geom_vline(xintercept = 23.16972)

```




```{r coeff-plot}
ggplot(data = county_slope, aes(x = county_plot, y = slope)) + 
	geom_point() +
    coord_flip() + 
	theme_bw() + 
	ylab("Change in hate crimes per year")  +
	xlab("County")
```




```{r bayesian-persistence-model}
# other option: Bayesian hierarchical model. 
install.packages("brms") # Hamiltonian monte carlo.
library(brms)
```






# Spatial Regression
Abandonment Age predicted by:
# Regression Analysis

Question: what factors best predict a) which pixels of agricultural land are abandoned, b) how long that land will stay abandoned for, and relatedly, c) how likely it is to be recultivated?

Factors:
- Climate variables (temperature, precipitation, growing degree days, etc.)
    - CHELSA (Swiss Federal Institute for Forest, Snow and Landscape Research WSL) ~1 km, 1979-2013, [CHELSA website](https://chelsa-climate.org/), [Peer reviewed paper, Scientific Data, 2017](https://www.nature.com/articles/sdata2017122)
    - Worldclim, ~1km, 1970-2000 - [website](https://www.worldclim.org/data/worldclim21.html), [bioclim](https://www.worldclim.org/data/bioclim.html),  [possibly helpful how-to](https://www.benjaminbell.co.uk/2018/01/extracting-data-and-making-climate-maps.html)
    

- Topography variables (elevation, slope, etc.) [earthenv.org](http://www.earthenv.org/topography)
- Agricultural economic variables 
    - soils
    - "marginality"
    - agricultural opportunity cost (Naidoo & Iwamura 2007)
    
- Socioeconomic variables
    - population
    - context
    - land use decision making
- 

## Methods:
- start with a regular OLS regression (lm)
- check out the residuals (the model errors) - if they are clumped, or there is a clear trend in them, then, they might be spatially autocorrelated.
- Calculating Moran's I helps identify spatial autocorrelation. 
- see code Liang shared by Raymond Huey 2020 ("/Users/christophercrawford/Google Drive/_Projects/abandonment_trajectories/scripts/Spatial analysis w Depth 2020-03-20.R")
- see code from Umesh's stats course. 



Citations:

## Code:


```{r load-data}

# Land use class codes:
#       1. Non-vegetated area (e.g. water, urban, barren land)
#       2. Woody vegetation
#       3. Cropland 
#       4. Herbaceous land (e.g. grassland)

# -------------------------- rasters --------------------------- #
# raw rasters
s <- brick(paste0(p_dat, "Abandonment/shaanxi.tif"))
b <- brick(paste0(p_dat_derived, "belarus.tif")) # merged version

# reprojected rasters:
br <- brick(paste0(p_dat_derived, "belarus_reproj.tif"))

# subset:
bt <- brick(paste0(p_dat_derived, "belarus_subset.tif"))

# age
s_age_r <- brick(paste0(p_dat_derived, "shaanxi_age.tif"))
b_age_r <- brick(paste0(p_dat_derived, "belarus_age.tif"))

# max_length
s_max_length_r <- brick(paste0(p_dat_derived, "shaanxi_max_length.tif"))
b_max_length_r <- brick(paste0(p_dat_derived, "belarus_max_length.tif"))

# update year names 1987 - 2017
names(s) <- paste0("y", 1987:2017)
names(b) <- paste0("y", 1987:2017)
names(br) <- paste0("y", 1987:2017)
names(s_age_r) <- paste0("y", 1987:2017)
names(b_age_r) <- paste0("y", 1987:2017)






# -------------------------- data.tables --------------------------- #

b_age <- fread(input = paste0(p_dat_derived, "belarus_age.csv"))
names(b_age)
s_age <- fread(input = paste0(p_dat_derived, "shaanxi_age.csv"))


b_length <- fread(input = paste0(p_dat_derived, "belarus_length.csv"))
s_length <- fread(input = paste0(p_dat_derived, "shaanxi_length.csv"))

b_max_length <- fread(input = paste0(p_dat_derived, "belarus_max_length.csv"))
s_max_length <- fread(input = paste0(p_dat_derived, "shaanxi_max_length.csv"))

# original data
s_dt <- fread(input = paste0(p_dat_derived, "shaanxi.csv"))
names(s_dt) <- gsub(pattern = "andcover", replacement = "y", names(s_dt))

b_dt <- fread(input = paste0(p_dat_derived, "belarus.csv")) # caution - huge file! 8.4 GB at least.
```

```{r simple-plots}
# crop b to a small area for testing
plot(bt$y2017, main = "Belarus 2017, subset", 
     col = plot_cols$color,
     breaks = c(0, plot_cols$breaks))
legend("bottomleft", cex = 0.7, inset = 0,
       legend = plot_cols$name, 
       fill = plot_cols$color)
```






```{r load-predictor-datasets}
# elevation
# slope
# soils

s_dt
s_age

```



