---
title: "Spatial Predictors of Abandonment and Recultivation"
author: "Christopher L. Crawford"
date: "2/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
bibliography: [/Users/christophercrawford/Google Drive/Library/library.bib]
---
# Spatial predictors of abandonment length and recultivation

*Need: read through papers, sort out my contribution.*
Insert context based on papers by @Estel2015 (recultivation, timing), @Estel2018, @Dara2018 (timing, recultivation, northern Kazakhstan), @Alcantara2013, @Lesiv2018 (methods of mapping), @Smaliychuk2016 (recultivation, predictors, Ukraine), @Prishchepov2013 (predictors), @Baumann2011 (), @Pazur2020 (Slovakia, predictors), @Levers2018a (predictors), @PerpinaCastillo2021 (predictors, model), @Meyfroidt2016 (drivers, tradeoffs, recultivation, timing, Russia/Ukraine/Kazakhstan), @Crouzeilles2020 (Atlantic Forest, predictors)



Question 3: what factors best predict abandonment trajectories?

a. What predicts which pieces of land are abandoned for the longest periods of time? In other words, are some areas more likely to experience more durable abandonment?
    i. Predicting the *max age* of a pixel [or *age in 2017*] using spatial predictor variables like population, slope, elevation, soil, etc.

b. What predicts recultivation? Are less suitable lands recultivated more quickly? Are more recently abandoned lands more frequently recultivated? In other words, does the probability that a piece of abandoned land will be recultivated depend on how long it has been abandoned for?
    i. My sense, from the abandonment decay plots, is that the longer a piece of land is abandoned for (i.e. the greater the greater the *age*), the less likely it is to be recultivated. This will involve predicting recultivation with factors like current age of abandonment, and the other variables included in the primary regression, etc. How will I signify "recultivation" in the dataset? Options include:
        1. data for each transition, each transition from abandoned -> not_abandoned is pulled out, and the age is recorded. This will require some fancy DT wrangling.
        2. all periods of abandonment, including instances when abandoned land remains abandoned (and the age it is), with the cohort, the year, the age, etc.). Each year after a piece of land is defined as abandoned. So, let's say I have a pixel that is in 1995 and stays abandoned for a total of 12 years. This would result in 9 total observations.

c. Taking two steps back - what predicts which pixels of agricultural land are abandoned?

```{r, echo = FALSE}
example_df <- data.frame(cohort = 1995,
           year = 2000:2012,
           age = 5:17,
           abn = c(rep("yes", 8), rep("no", 5)),
           recultivated = c(rep("no", 8), rep("yes", 5)))

example_df[1:9, ]
```

## Predictor variables:

1. Agricultural suitability
    a. *Growing Degree Days* (the cumulative number of degrees above growing threshold accumulated over the course of a growing season. 18 - 10 [threshold] * the number of days at that temp.). See [the University of Wisconsin's website](https://nelson.wisc.edu/sage/data-and-models/atlas/maps.php?datasetid=31&includerelatedlinks=1&dataset=31) for a good description of how GDD works. I plan to make use of GDD above base of 5 degrees C, [from the ENVIREM dataset](http://envirem.github.io/).  
    b. *Soil Quality*: 
        + [ISRIC World Soil Grids](https://www.isric.org/explore/soilgrids), at a resolution of 250 m. Used by @Levers2018a.
        + [Harmonized World Soil Database](https://daac.ornl.gov/SOILS/guides/HWSD.html), which has a resolution of ~ 5.5 km (0.05 degrees). (I plan to use soil organic carbon, soil pH, and soil fertility)
    c. FAO's Global Agro-Ecological Zones (GAEZ), which includes both general natural resource, soil, terrain, slope, elevation, and other types of biophysical data, typically at the scale of 5 arc minutes (~10km at the equator), as well as crop specific suitability maps (e.g. for rain-fed winter wheat). Specifically:
        i. Agro-ecological zones (categories such as: steep terrain, dry/good soils, dry/poor soils, sub-humid/good soils...)
        ii. Soil types
        iii. Workability (categories such as: no constraints, moderate constraints, severe constraints...)
        iv. And more.
    d. [Global inherent land quality map](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/use/worldsoils/?cid=nrcs142p2_054029)
    e. Agricultural opportunity cost (Naidoo & Iwamura 2007)
    f. Travel time to major cites - [Year 2000](https://forobs.jrc.ec.europa.eu/products/gam/), [Year 2015](https://map.ox.ac.uk/research-project/accessibility_to_cities/)
2. Environmental factors
    a. Temperature and precipitation (*use GDD, see above, as a nice*). [Possibly helpful how-to](https://www.benjaminbell.co.uk/2018/01/extracting-data-and-making-climate-maps.html)
        i. [Terraclimate](http://www.climatologylab.org/terraclimate.html)
        ii. [Bioclim](https://www.worldclim.org) - ~1km, 1970-2000, [website](https://www.worldclim.org/data/worldclim21.html),
        iii. CHELSA (Swiss Federal Institute for Forest, Snow and Landscape Research WSL) ~1 km, 1979-2013, [CHELSA website](https://chelsa-climate.org/), [Peer reviewed paper, Scientific Data, 2017](https://www.nature.com/articles/sdata2017122)
        iv. Climate Hazards group Infrared Precipitation with States ([CHIRPS](https://www.nature.com/articles/sdata201566))
    b. *Slope, elevation, ruggedness*, and other terrain variables:
        + see what is on GEE!
        + *Ruggedness (Terrain Ruggedness Index)* is a metric of elevation differences between neighboring cells, first derived by @Riley1999. It calculates the "the difference in elevation values from a center cell and eight cells directly surrounding it. Then, the eight elevation differences are squared and averaged. The square root of this average results is a TRI measurement for the center cell" ([see more here](https://community.esri.com/t5/water-resources-blog/terrain-ruggedness-index-tri-and-vector-ruggedness-measurement/ba-p/884340)). Other people use the "vector ruggedness measure," one of the variables supplied by @Amatulli2018.
        + @Amatulli2018 (~1 km, [see earthenv.org](http://www.earthenv.org/topography)), which has a lot of different variables at 1km (elevation, slope, aspect, eastness, northness, roughness, terrain roughness index, topographic position index, vector ruggedness measure)
        + SRTM ~90m, 3 arc seconds (~ 3 / 60 / 60 * 110 * 1000 = `r 3/60/60 * 110 * 1000` at equator) from [EarthEnv](http://www.earthenv.org/DEM)
        + 3 arc second DEM from Jonathan de Ferranti (de Ferranti, J. Digital Elevation Data. Viewfinder Panoramas http://www.viewfinderPanoramas.org/dem3.html (2015). Used in development of WorldPop.
    c. Surrounding landcover, i.e. proximity to woody veg/grassland
        + Distance to grassland
        + Distance to forest
        + Proportion of pixel that is abandoned. If the pixels are aggregated to a coarser scale, from 30m x 30m to 1km x 1km ($1 km^2$), this is the proportion of pixels within that $1 km^2$ that were abandoned at all. This is similar to "Distance to permanently fallowed areas" variable used by @Levers2018a. 
        + Distance to pixels that had experienced abandonment at any point prior to that year.
        + 
3. Socioeconomic variables:
    a. Population, either from SEDAC (GPW, GRUMP, etc.), WorldPop, or [UN DESA’s World Urbanization Prospects](https://esa.un.org/unpd/wup/)
        + Population rates of change
        + Proportion of population in elderly age class (or average age of population)
        + Population in working age
        + Birth rates
    c. Land use decision making, @Malek2020. [download dataset here](https://dataverse.nl/dataset.xhtml?persistentId=doi:10.34894/JEDNM5). 10 km resolution.
4. Farm management factors:
    a. Field size, from Fritz et al. 2015. This was found by @Levers2018a to be a significant factor (though not the most significant)

@Levers2018a explored the following predictors:
1. Area equipped for irrigation
2. Field size.
3. Average livestock density
4. Average yields from grain crops
5. Aridity index (mean annual precip. / mean annual potential evapotranspiration)
6. Growing degree days
7. Soil cation exchange (soil fertility)
8. Distance to forested areas
9. Distance to permanent fallow areas (i.e. areas that were fallow for most of the time series between 2001-2012)
10. Soil organic carbon (overall soil health).
11. Soil pH value.
12. Terrain ruggedness
13. Access - travel time to cities with more than 50,000 inhabitants
14. Population density
15. Change in population density from 2001-2006, 2007-2012. 
16. Average unemployment ratio.

@Levers2018a also used a method that allowed the coefficient on each predictor to vary spatially, meaning that something like field size could have a positive effect in the UK, but a negative one in the Alps. In my case, I'll allow for this by allowing for site-level fixed effects. 

@Levers2018a found that the factors that seemed to be the best predictors were (in order):
1. Growing degree days (which varied through space).
2. Field size (which strongly varied through space)
3. Distance to permanently fallowed land.
4. Unemployment
5. Soil pH
6. Soil organic carbon
7. Change in population density
8. Yields.





### Population data sources

SEDAC - Columbia University's CIESIN (Center for International Earth Science Information Network)  
* Gridded Population of the World (GPW v4.11) years *2000, 2005, 2010, 2015, 2020*, ~ 1 km resolution, for 
  + [population density (# people per km$^2$)](https://beta.sedac.ciesin.columbia.edu/data/set/gpw-v4-population-density-rev11). 
  + [population counts](https://beta.sedac.ciesin.columbia.edu/data/set/gpw-v4-population-count-rev11)
  + [density adjusted to match UN country totals](https://beta.sedac.ciesin.columbia.edu/data/set/gpw-v4-population-density-adjusted-to-2015-unwpp-country-totals-rev11)
  + [counts adj to UN](https://beta.sedac.ciesin.columbia.edu/data/set/gpw-v4-population-count-adjusted-to-2015-unwpp-country-totals-rev11)
  + [see collection overview (this is SEDAC's flagship data product)](https://beta.sedac.ciesin.columbia.edu/data/collection/gpw-v4)

* Global Rural-Urban Mapping Project (GRUMP) v1, years *1990, 1995, 2000*, at ~ 1 km resolution
  + [population density (population per km$^2$)](https://sedac.ciesin.columbia.edu/data/set/grump-v1-population-density)
  + [population counts](https://beta.sedac.ciesin.columbia.edu/data/set/grump-v1-population-count)

* Global Population Density Grid Time Series back-casts: *1970, 1980, 1990, 2000*, ~ 1km
  + [population density](https://sedac.ciesin.columbia.edu/data/set/popdynamics-global-pop-density-time-series-estimates)
  + [population counts](https://sedac.ciesin.columbia.edu/data/set/popdynamics-global-pop-count-time-series-estimates)

* Global Estimate Net Migration Grids, for *1970, 1980, 1990, 2000*, ~ 1km
  + [download here](https://sedac.ciesin.columbia.edu/data/set/popdynamics-global-est-net-migration-grids-1970-2000)

* SSP Future projections: decadal estimates or projections of urban, rural, and total population at ten-year intervals between *2000-2100* [download from SEDAC here](https://beta.sedac.ciesin.columbia.edu/data/set/popdynamics-1-km-downscaled-pop-base-year-projection-ssp-2000-2100-rev01), based on @Jones2016.


UN DESA  
- future projections of population growth and urbanization from UN DESA’s World Urbanization Prospects [see here](https://esa.un.org/unpd/wup/). Note, this is not a raster, just a spreadsheet with country totals.


WorldPop: high resolution (~100m, 1 km), annually from *2000-2020* - [download at WorldPop.org](https://www.worldpop.org/project/categories?id=3), [scientific data paper](https://www.nature.com/articles/sdata20171).  
* Download individual countries, at either 100m or 1km resolution.
* Like GPW, they have raw counts as well as those that have been adjusted to match UN population estimates from 2019.
* Note: "constrained" vs. "unconstrained" just refers to whether the top-down datasets have been constrained based on satellite image derived maps of the built environment. This data is only available for the year 2020, so I'm going to use the "unconstrained top-down" dataset so that I can use multiple years.
* Top-down vs. bottom-up refers to whether estimates are downscaled from accurate national statistics onto a grid (top-down), or derived from local estimates at particular sample locations scaled up to a grid (bottom-up). I'm using the "top-down" datasets.

#### Pros and Cons:
* WorldPop may be more accurate than GPW, [at least in places like China](https://link.springer.com/article/10.1007/s11111-020-00366-2). However, it only exists for 2000-2020. 
* On the other hand, SEDAC's datasets (GPW/GRUMP) cover the full period from 1970-2020, with consistency across datasets that allows for comparison. 
* Decision: due to the fact that WorldPop only covers the time period from 2000-2020, I will use SEDAC's datasets. 


Other:  
- Human Footprint maybe? No - just use the direct population density.


### My chosen predictors




## Results outline

3. Spatial predictors of age and recultivation

- coefficients on various predictor variables
- map of *abandonment age in 2017*
- map of *max age*

Fig. 8. Coefficients on spatial predictors (vertical, showing coefficients on slope, elevation, suitability, population)
Fig. 9. Maps of a) abandonment *age in 2017* and b) *max age* for example site.




```{r start}
source("scripts/0_start.R")
env_size(ls())

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-site-data}
# copied from 6_figures on Feb 1 2021

# for use on cluster:
# p_dat <- "/scratch/network/clc6/abandonment_trajectories/data_derived/"
# p_dat_derived <- "/scratch/network/clc6/abandonment_trajectories/data_derived/"
# p_output <- paste0("/scratch/network/clc6/abandonment_trajectories/output/")
list.files(p_dat_derived)

site_df <- read.csv(file = paste0(p_dat_derived, "site_df.csv"))


# Land use class codes:
#       1. Non-vegetated area (e.g. water, urban, barren land)
#       2. Woody vegetation
#       3. Cropland 
#       4. Herbaceous land (e.g. grassland)

# testers:
bs <- brick(paste0(p_dat, "Abandonment/belarus_small.tif"))
bt <- brick(paste0(p_dat_derived, "belarus_subset.tif"))
names(bs) <- paste0("y", 1987:2017)
names(bt) <- paste0("y", 1987:2017)


# -------------------------- rasters --------------------------- #
# raw rasters
s <- brick(paste0(p_dat_derived, "input_rasters/shaanxi.tif"))
b <- brick(paste0(p_dat_derived, "input_rasters/belarus.tif")) # merged version

# age
s_age_r <- brick(paste0(p_dat_derived, "shaanxi_age.tif"))
b_age_r <- brick(paste0(p_dat_derived, "belarus_age.tif"))

# max_length
s_max_length_r <- brick(paste0(p_dat_derived, "shaanxi_max_length.tif"))
b_max_length_r <- brick(paste0(p_dat_derived, "belarus_max_length.tif"))

# update year names 1987 - 2017
names(s) <- paste0("y", 1987:2017)
names(b) <- paste0("y", 1987:2017)
names(s_age_r) <- paste0("y", 1987:2017)
names(b_age_r) <- paste0("y", 1987:2017)


# prepared input rasters (derived by Chris)
site_input_raster_files <- list.files(paste0(p_dat_derived, "input_rasters"), full.names = TRUE) %>%
  grep(".tif", ., value = TRUE) #%>% grep("age", ., value = TRUE, invert = TRUE)

r <- lapply(seq_along(site_input_raster_files), function(i) {brick(site_input_raster_files[i])})
names(r) <- site_df$site

# rename raster layers:
for (i in 1:11) {
  if (names(r[i]) == "nebraska") {
    names(r[[i]]) <- paste0("y", 1986:2018)
  } else {
    if (names(r[i]) == "wisconsin") {
      names(r[[i]]) <- paste0("y", 1987:2018)
    } else {
      # everything else, just 1987:2017
      names(r[[i]]) <- paste0("y", 1987:2017)
    }}}


# abandonment age maps (produced by Chris)
age_files <- list.files(paste0(p_dat_derived, "age_rasters"), full.names = TRUE) %>%
  grep(".tif", ., value = TRUE) #%>% grep("age", ., value = TRUE, invert = FALSE)

age_r <- lapply(seq_along(age_files), function(i) {brick(age_files[i])})
names(age_r) <- site_df$site
for (i in seq_along(age_r)) {names(age_r[[i]]) <- paste0("y", 1987:2017)} # remember: these are just 1987:2017

# year of first abandonment maps (from He)
yoa_files <- list.files(paste0(p_dat, "Abandonment/year_of_abandonment/"))

# -------------------------- data.tables --------------------------- #
b_age <- fread(input = paste0(p_dat_derived, "belarus_age.csv"))
names(b_age)
s_age <- fread(input = paste0(p_dat_derived, "shaanxi_age.csv"))


b_length <- fread(input = paste0(p_dat_derived, "lengths/", "belarus_length_b1.csv"))
s_length <- fread(input = paste0(p_dat_derived, "lengths/", "shaanxi_length_b1.csv"))

b_max_length <- fread(input = paste0(p_dat_derived, "lengths/", "belarus_max_length_b1.csv"))
s_max_length <- fread(input = paste0(p_dat_derived, "lengths/", "shaanxi_max_length_b1.csv"))

# original data
s_dt <- fread(input = paste0(p_dat_derived, "shaanxi.csv"))
names(s_dt) <- gsub(pattern = "andcover", replacement = "y", names(s_dt))

b_dt <- fread(input = paste0(p_dat_derived, "belarus.csv")) # caution - huge file! 8.4 GB at least. 



# -------------------------- summarized data.frames --------------------------- #
indx <- 9
site <- site_df$site[indx] # set site:
site_label <- site_df$label[indx] # set label
blip_label <- "_b1"
load(file = paste0(p_output, "abn_dat_products", blip_label, site_label, ".rds"), verbose = TRUE)

# loads:
area_b1_s
persistence_list_b1_s
abn_area_change_b1_s



```

```{r load-predictor-data}
# elevation
# de Ferranti's data:
# 110 - 112 E
# 38 - 40 N


# slope
# soils

```

```{r simple-plots}
# copied from 6_figures on Feb 1 2021

# -------------------------- plot raw data --------------------------- #
plot(s$y1987, main = "Sha(a)nxi 1987", breaks = c(0, plot_cols$breaks), col = plot_cols$color)
legend("bottomleft", cex = 0.6, inset = 0,
       legend = plot_cols$name, 
       fill = plot_cols$color)

plot(s$y2017, main = "Sha(a)nxi 2017", breaks = c(0, plot_cols$breaks), col = plot_cols$color)
legend("bottomleft", cex = 0.6, inset = 0,
       legend = plot_cols$name, 
       fill = plot_cols$color)

# ---------------------- plot abandonment age ---------------------- #
plot(s_age_r$y2017, main = "Sha(a)nxi Age of Abandonment 2017")
plot(s_age_r[[28:31]])

# -------------------------- animate --------------------------- #
# --------------
# https://www.rdocumentation.org/packages/raster/versions/3.1-5/topics/animate

animate(bt, pause = 0.5, zlim = c(1, 4), maxpixels=5000, n=1,
        breaks = c(0, plot_cols$breaks), col = plot_cols$color)

```

## Methods

- start with a regular OLS regression (lm)
- check out the residuals (the model errors) - if they are clumped, or there is a clear trend in them, then, they might be spatially auto-correlated.
- Calculating Moran's I helps identify spatial autocorrelation. 
- see code Liang shared by Raymond Huey 2020 ("/Users/christophercrawford/Google Drive/_Projects/abandonment_trajectories/scripts/Spatial analysis w Depth 2020-03-20.R")
- see code from Umesh's stats course. 

@Levers2018a use the R package `mboost` from [Hothorn et al. 201o](http://mboost.r-forge.r-project.org/) to implement "boosted regression trees" or BRT. This method has an advantage because, as quoted from @Pazur2020, "The advantages of BRT modelling compared to conventional statistical methods, such as ordinary least squares and logistic regressions, is that it allows capturing the non-linear relationship between the dependent and independent variables, does not require a normal distribution assumption and is robust against the multicollinearity of variables (Elith et al. 2008)." 

I had initially been thinking that I wouldn't need to account for too much spatial variation, since the models can have site fixed effects that account for this spatial variation. But, it might be worth exploring these BRT models.

See also the package `gbm`, and the documentation from the package [`dismo`](https://rspatial.org/raster/sdm/9_sdm_brt.html).

@Dara2018

@Pazur2020 and @Pazur2020a also use boosted regression trees

```{r}

```

In order to perform a regression, I need to aggregate abandonment age pixels to match the coarser scale of my predictors (most of which are ~1km)
```{r aggregate-abn-age}
# plot
# -------------------------- plot raw data --------------------------- #
plot(s$y1987, main = "Sha(a)nxi 1987", breaks = c(0, plot_cols$breaks), col = plot_cols$color)
legend("bottomleft", cex = 0.6, inset = 0,
       legend = plot_cols$name, 
       fill = plot_cols$color)

plot(s$y2017, main = "Sha(a)nxi 2017", breaks = c(0, plot_cols$breaks), col = plot_cols$color)
legend("bottomleft", cex = 0.6, inset = 0,
       legend = plot_cols$name, 
       fill = plot_cols$color)

# ---------------------- plot abandonment age ---------------------- #
plot(s_age_r$y2017, main = "Sha(a)nxi Age of Abandonment 2017")
plot(s_age_r[[28:31]])



# 30 m resolution -> 1000 m resolution ------------------------------------------- #
# In order to go from 30 m to 1000m, first I downscale the 30 m grid to a 10 m grid, 
# by disaggregating with a factor of 3. This yields nine 10 m x 10 m pixels within each 
# of the original pixels. These are then aggregated by factor of 100, in order to 
# get to a 1000 m x 1000 m grid.

(1 / (3^2)) * (30/1000)^2 # 1/10,000
(1/100)^2

(1/3)^2 * (100)^2 # There are 1111.11 30m x 30m pixels in a 1km x 1km pixel.


ncell(s_age_r$y2017)

# create intermediary raster, to aggregate 
sa <- s_age_r$y2017


# age raster, with NA values set to 0:
sa_na_0 <- s_age_r$y2017
sa_na_0[is.na(sa_na_0)] <- 0

# all the same:
plot(sa, main = "sa, including NA")
plot(sa_na_0, main = "sa, with NA set to 0")
plot(s_age_r$y2017, main = "(a) Shaanxi: abandonment age in 2017")


# aggregate pixels to 1km x 1km.
sa_agg_mean <- aggregate(
  disaggregate(sa, fact = 3), 
  fact = 100, fun = mean)

sa_agg_mean_na_0 <- aggregate(
  disaggregate(sa_na_0, fact = 3), 
  fact = 100, fun = mean)

plot(sa_agg_mean, main = "(b) Aggregated with mean, NA removed")
plot(sa_agg_mean_na_0, main = "(c) Aggregated with mean, NA set to 0")


# create a binary abandonment raster
sa_bin <- sa
sa_bin[!is.na(sa_bin)] <- 1
plot(sa_bin, main = "Binary abandonment map (2017)")

# calculate the proportion of 30m x 30m pixels within 1km x 1km pixel
# that are abandoned:
sa_count <- aggregate(
  disaggregate(sa_bin, fact = 3), 
  fact = 100, fun = sum) / (100^2) # note that because the count is of 10 m pixels (there are 9 of these in each 30m pixel), this must be divided by 100^2

plot(sa_count, main = "(d) Proportion of pixel abandoned")


# note that this is the same as mean with NAs set to 0
plot(sa_agg_mean*sa_count, main = "Weighted average age")
plot(sa_agg_mean_na_0, main = "(c) Aggregated with mean, NA set to 0")


# other aggregation functions ----------------------------------------------- #

sa_agg_max <- aggregate(
  disaggregate(sa, fact = 3), 
  fact = 100, fun = max)

sa_agg_median <- aggregate(
  disaggregate(sa, fact = 3), 
  fact = 100, fun = median)

plot(sa_agg_max, main = "aggregation method: max")
plot(sa_agg_median, main = "aggregation method: median")
```

```{r age-histogram}
s_age <- fread(input = paste0(p_dat_derived, "shaanxi_age.csv"))


# histogram of age values in 2017
s_age[, .N, by = y2017][order(y2017)][y2017 >= 5] %>% data.frame() %>% 
  ggplot(data = ., mapping = aes(x = y2017, y = N)) + 
  geom_col()


# hist(s_age_r$y2017) # Caution: this is very slow. 
```


```{r save-agg-rasters-to-png}
# First, comparing the aggregation methods: decided to use the mean.
png(file = paste0(p_output, "plots/age_aggregation_comparison.png"),
      width = 10, height = 3, units = "in", res = 400)
par(mfrow = c(1, 4), oma = c(0,0,0,1))

plot(s_age_r$y2017, main = "Shaanxi Age, 2017")
plot(sa_agg_mean, main = "Aggregated w mean")
plot(sa_agg_max, main = "Aggregated w max")
plot(sa_agg_median, main = "Aggregated w median")
dev.off()




# Second, comparing the NA removal methods: decided to leave the NAs removed 

png(file = paste0(p_output, "plots/age_agg_na_rm_v_na_0_wide.png"),
    width = 11, height = 3,
    units = "in", res = 400)
par(mfrow = c(1, 4), oma = c(0,0,0,1))
plot(s_age_r$y2017, main = "(a) Shaanxi: abandonment age in 2017")
plot(sa_agg_mean, main = "(b) Aggregated with mean, NA removed")
plot(sa_agg_mean_na_0, main = "(c) Aggregated with mean, NA set to 0")
plot(sa_count, main = "(d) Proportion of pixel abandoned")
dev.off()

```


```{r regression-methods}

```


