---
title: "The impermanence of cropland abandonment limits potential for environmental benefits (draft)"
author: Christopher L. Crawford,$^a$$^*$ He Yin,$^b$ Volker Radeloff,$^c$ and David S. Wilcove$^{a, d}$
date: "`r format(Sys.time(), '%B %d, %Y')`"
abstract: |
  | $^a$Princeton School of Public and International Affairs, Princeton University, Princeton, NJ
  | $^b$Department of Geography, Kent State University, Kent, OH
  | $^c$Department of Forest & Wildlife Ecology, University of Wisconsin-Madison, Madison, WI
  | $^d$Department of Ecology & Evolutionary Biology, Princeton University, Princeton, NJ
  |
  | $^*$Corresponding Author, ccrawford@princeton.edu, Robertson Hall, Princeton University, Princeton, NJ
output:
  bookdown::word_document2:
    reference_docx: /Users/christophercrawford/Google Drive/_Projects/writing/scripts/word_style_ref.docx
    number_sections: false
  bookdown::pdf_document2: 
    keep_tex: true
  bookdown::html_document2:
    toc: true
editor_options:
  chunk_output_type: console
bibliography: [/Users/christophercrawford/Google Drive/Library/library.bib, packages.bib]

---

```{r set-chunk-opts, include = FALSE}
knitr::opts_chunk$set(
  #out.width = "1\\textwidth",  # if compiling a pdf, include this option
  #comment=NA #fig.width=6, fig.height=6
  echo = FALSE
  )

# nocite: |
#   @RStudioTeam2020, @R-rmarkdown, @R-bookdown, @R-knitr
```
```{r load-packages-functions, eval = TRUE, echo = FALSE, include = FALSE}
source("/Users/christophercrawford/Google Drive/_Projects/abandonment_trajectories/scripts/0_start.R")
source(paste0(p_proj, "scripts/util/_util_master.R"))
```


```{r load-files, include = FALSE, eval = TRUE}
# set run:
run_label <- "_2021_03_13"

# Additional paths --------------------------------------------------- #
p_plots <- "/Users/christophercrawford/Google Drive/_Projects/abandonment_trajectories/output/plots/"

# load general spatial data on site locations --------------------------------------------------- #
site_df <- read.csv(file = paste0(p_dat_derived, "site_df.csv"))
site_sf <- st_read(paste0(p_dat_derived, "site_sf.shp"))



# load all length data (distilled)
length_distill_df <- read_csv(file = paste0(p_dat_derived, run_label, "/", 
                                            "length_distill_df", run_label, ".csv"))

# load mean length data
mean_length_df <- read_csv(file = paste0(p_dat_derived, run_label, "/", "mean_length_df", run_label, ".csv"))

# load summary statistic derived from lengths
summary_stats_all_sites <- read_csv(file = paste0(p_dat_derived, run_label, "/", "summary_stats_all_siteS", run_label, ".csv"))
summary_stats_all_sites_pooled <- read_csv(file = paste0(p_dat_derived, run_label, "/", "summary_stats_all_sites_pooled", run_label, ".csv"))


# load area summary stats:
area_summary_df <- read_csv(file = paste0(p_dat_derived, run_label, "/", "area_summary_df", run_label, ".csv"))



# load combined persistence, area, and turnover datasets:
persistence_dat <- read_csv(file = paste0(p_dat_derived, "abn_decay_data", run_label, ".csv")) %>%
  mutate(site = as.factor(site),
         bins = as.factor(bins),
         year_abn_bins = as.factor(year_abn_bins),
         cohort = as.factor(cohort))

area_dat <- read_csv(file = paste0(p_dat_derived, "area_dat", run_label, ".csv"))
turnover_dat <- read_csv(file = paste0(p_dat_derived, "turnover_dat", run_label, ".csv"))



# load decay rate data
load(file = paste0(p_dat_derived, run_label, "/decay_mod_coef_dfs.rds"), 
     verbose = TRUE)

# load extrapolation data
extrapolate_df_l3 <- read_csv(file = paste0(p_dat_derived, run_label, "/extrapolate_df_l3", run_label, ".csv"))


```

```{r collect-stats, include = FALSE}
# Testing how the tibble looks `r filter(mean_length_df, abn_threshold == 5, length_type == "all") %>% ungroup() %>% slice_min(order_by = mean_duration) %>% select(site) %>% unlist() %>% capwords()`

filter(mean_length_df, abn_threshold == 5, length_type == "all") %>% ungroup() %>% summarise(min = min(mean_duration))

filter(mean_length_df, abn_threshold == 5, length_type == "all") %>% arrange(mean_duration) %>% .[1, ] %>% .$site

filter(mean_length_df, abn_threshold == 5, length_type == "all") %>%
  ungroup() %>% slice_min(order_by = mean_duration) %>% select(site) %>% unlist() %>% capwords()



mean_length_df %>% 
  filter(abn_threshold == 5, length_type == "all") %>% #select(mean_duration) %>% 
  arrange(mean_duration) #%>% summary()

filter(mean_length_df, abn_threshold == 5, length_type == "all") %>% .$mean_duration %>% min()
filter(mean_length_df, abn_threshold == 5, length_type == "all") %>% .$mean_duration %>% max()

summary_stats_all_sites %>% filter(abn_threshold == 5, length_type == "all") %>% .$mean_of_means


# decay mod numbers:

mean_coef_df_wide_l3 %>% filter(site == "shaanxi") %>% .$time_to_0.5 %>% round(digits = 1)


# extrapolation
extrapolate_df_l3 %>% select(site, max_mean_age) %>% unique() %>% arrange(max_mean_age) %>% .$max_mean_age %>% .[9] %>% round(digits = 2)

```

Last edited on April 22nd, 2021, copied to ms document.

# Abstract
*Must be <250 words (currently 229)*

Agricultural expansion is a major cause of land-use change globally, but millions of hectares of cropland are being abandoned as a result of demographic, economic, and environmental changes.
These abandoned croplands could be immensely valuable for carbon sequestration and biodiversity restoration.
However, their environmental value depends on the duration and persistence of abandonment, which is poorly known.
Here, we use new high-resolution satellite imagery to examine the timing and duration of abandonment at eleven sites across four continents (and various environments) that have experienced cropland abandonment from 1987 and 2017.
We find that abandonment is by and large fleeting, lasting on average only `r summary_stats_all_sites$mean_of_means %>% round(digits = 2)` years (SD = `r summary_stats_all_sites$mean_of_sds %>% round(digits = 2)`).^[*Note: both values reported are the means of each site's corresponding summary statistics, i.e. the mean of means and mean of standard deviations*]
At most sites, more than 50% of abandoned croplands were recultivated within 30 years, and some sites, particularly in Eastern Europe, Russia, and the US, showed accelerating rates of recultivation.
If current rates of abandonment continue, the mean duration of abandonment at most sites will likely plateau between `r extrapolate_df_l3 %>% select(site, max_mean_age) %>% unique() %>% arrange(max_mean_age) %>% .$max_mean_age %>% .[1] %>% round(digits = 1)` and `r extrapolate_df_l3 %>% select(site, max_mean_age) %>% unique() %>% arrange(max_mean_age) %>% .$max_mean_age %>% .[9] %>% round(digits = 1)` years by about 2040.
Counter to optimistic assumptions, most abandonment is unlikely to be permanent and will produce little in the way of carbon or biodiversity benefits.
New policies and incentives will be needed to lengthen the period of abandonment so as to generate these and other benefits.
Until then, abandoned croplands will remain untapped opportunities.

# Significance Statement
*Must be <120 words (currently 121)*

Demographic, economic, and environmental changes often result in cropland abandonment. 
While sometimes seen as a threat to food security and cultural landscapes, abandonment is increasing billed as an opportunity for habitat restoration and carbon sequestration.
However, those environmental benefits largely depend on how long ecosystems can regenerate in abandoned croplands and accumulate carbon and biodiversity, which takes decades in many ecosystems.
Existing studies often rely on single snapshots in time to estimate abandonment, and ignored recultivation.
Using annual land-cover maps, we track abandonment and recultivation at eleven sites across four continents that have recently experienced abandonment.
Contrary to previous assumptions, abandonment is short-lived and lands are frequently recultivated.
Until policymakers take steps to reduce recultivation, abandonment will remain a missed opportunity.


```{r Hélène's-suggestions}
# A few more specific comments, which you are very welcome to entirely ignore:
# On the abstract:
# - when you make the point that such opportunities depend on the duration and persistence of abandonment, does this hold for all alternatives describes just above, or especially for habitat restoration?
# - when you first mention the 11 sites, it would be helpful to say that they are all around the world, it's a big value added to your study I think. Although I noticed you're supposed to cut words __
# - you might want to make the "counter to recent assumptions" more prominent, it's a big part of your selling point I think __
# - the last sentence of both the abstract and the significance statement are really cool!!
# On the significance statement:
# - small detail: maybe remove one of the two "sometimes" in the second line 
# - same remark about the 11 sites, I think it would be helpful to add that they are all around the world
# - can you say that abandonment can be recultivated? If so, no need to change; it just sounded a bit strange
# - again, the closing sentence is great!
```


```{r cuts}
# 
# Only one site showed more durable abandonment: `r site_df$description[9]`, where projections require `r mean_coef_df_wide_l3 %>% filter(site == "shaanxi") %>% .$time_to_0.5 %>% round(digits = 1)` years for 50% abandoned cropland to be recultivated, likely a result of Chinese reforestation and afforestation programs.
```


```{r old-language, include = FALSE}
# ## old
# 
# Environmental benefits of regenerating old farmland 
# Unless 
# For environmental benefits in particular, timing is everything.  a result,  policymakers 
# 
# benefits of  for carbon sequestration and biodiversity 
# Given that biodiversity may recover slowly in both forest and grassland ecosystems, and carbon sequestration benefits only last as long as  carbon sequestration may 
# For environmental benefits in particular, timing is everything. 
# 
# abandonment has lacked strong empirical data on long-term outcomes.
# Many recent studies rely on two snapshots in time to produce estimates of abandonment. 
# Our research shows that these estimates may overestimate the durability of abandonment.
# Lacking a more detailed understanding of the impermanent nature of abandonment, . 
# Biodiversity may 
# 
# Some biomes may undergo succession more quickly than others, yet
# 
# which are required to inform both policies and research into potential opportunities for this land
# 
# To date, however, the understanding of abandonment has been limited by data availability.
# Maps of abandonment have been coarse in both spatial and temporal resolution, resulting in potentially misleading estimates of availability.
# 
# 
# ## alternative language
# 
# Recent calls for large-scale land restoration to address climate change and biodiversity loss have highlighted this reservoir of land for its potential to provide environmental benefits in the form of regenerating wildlife habitats and carbon sequestration. 
# However, the availability of this land for alternative uses, and any potential environmental benefits, depend first and foremost on the duration and durability of abandonment, and therefore how long habitats are allowed to regenerate.
# 
# Yet permanent abandonment is far from guaranteed: abandoned lands are increasingly sought after for increased cultivation, renewable energy, and biofuel production, calling the potential for long-term regeneration into question. 
# Using a time-series of high-resolution maps of abandonment, I ask the following: how long does abandoned agricultural land typically remain abandoned for, and how does this vary geographically across sites? Then, by modeling the recultivation of abandoned lands through time, I provide insights about the potential benefits that may accrue into the future if current recultivation trends continue. Finally, I investigate which spatial factors best explain which lands are abandoned for longer periods of time. Together, these results provide a clearer picture of the potential for habitat regeneration to take place in abandoned croplands, and in turn, to provide benefits in terms of species habitat gains and carbon sequestration. As abandonment trends continue around the world, these results will help improve policy responses to abandonment, either to reduce frequent turnover and gain more benefits from abandoned lands, or to target conservation investment and protection to those areas that are more likely to provide environmental benefits.
```
