---
title: "Abandonment Decay Model Explorations"
author: "Christopher Crawford, Princeton University"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---
This document is designed to should do the following:
1. Load the packages you need and the data on abandonment decay (I label “dat_l" in the script). This chunk also includes descriptions of each of the variables.
2. Run the current iteration of the linear model, with the log log structure and cohort as a fixed effect, for just one site out of eleven (Shaanxi, in this example)
3. Use lapply() to run five versions of the model across all 11 sites.
4. Plot the results of just one site (Shaanxi, in this example). Note the variables on the x-axis. There are two ways to plot this (with age, which starts at 5, or with time_1, which starts at 1. 
5. Plot and save the diagnostic plots for all 11 sites, first sequentially and then the fitted v. Residuals plot in one mondo figure.
6. Run a single lm for all sites, with the log-log structure and cohort and now site as fixed effects.
7. Try a model including a log term for time as well as a linear term for time.
8. Plot residuals vs. covariates.

```{r packages, include = FALSE}
# packages_list <- c("tidyverse")
library(tidyverse)
library(tidymodels)
```


```{r load-data, include = FALSE}
# update this path to wherever you want to keep the data, output plots, etc.
your_path <- "/Users/christophercrawford/Google Drive/_Projects/abandonment_trajectories/data_derived/" # 


site_names <- c("belarus", "bosnia_herzegovina", "chongqing", "goias", "iraq", "mato_grosso", "nebraska", "orenburg", "shaanxi", "volgograd", "wisconsin")

# load the data
dat_l <- read_csv(file = paste0(your_path, "abn_decay_data.csv"))

# restructure factors and such
dat_l <- mutate(dat_l, 
                site = as.factor(site),
                bins = as.factor(bins),
                year_abn_bins = as.factor(year_abn_bins),
                cohort = as.factor(cohort))


# select just a subset of this, for model predictions with augment() - this is only really used for a few things
dat_l_0 <- select(dat_l, site, time_1, age, initial_area_abn, cohort, year_abn)


str(dat_l)
# variables:
# variables of interest

dat_l$site %>% unique() # 11 sites in the dataset.

dat_l$proportion # response variable: the proportion of originally abandoned land that remains in a given year. 
dat_l$year # the current year, allowing us to measure area abandoned in the current year, relative to the amount initially abandoned
dat_l$year_abn # cohort of land abandoned in a particular year, which is then followed year after year to track persistence
# ^ probably should be a factor
dat_l$count # number of pixels, from which area is determined
dat_l$area_ha # area abandoned, of a given cohort
dat_l$age # how long that land in that cohort has been abandoned for

dat_l$time # age - 5; adjusted time abandoned beyond the initial abandonment threshold. So, 0 corresponds to an abandonment age of 5, 1 -> 6, 2 -> 7 etc.

dat_l$time_1 # age - 4 (i.e. time + 1). This is so that time starts from 1, to avoid issues with taking the log of 0. 

dat_l$bins # bins of abandoned land of similar ages (5-10 years)
dat_l$year_abn_bins # bins of land abandoned during the same five to six year window
dat_l$initial_area_abn # the area (in ha) initially abandoned in a given year (cohort)

dat_l$cohort # a factor of "year_abn" to be used for cohort 
```


# Testing for just one site:
```{r lm-log-log-cohort}
# log-log, with cohort fixed effects: for just one site

lm_mod_log2_cohort <- lm(log(proportion) ~ 0 + log(time_1):cohort, data = filter(dat_l, site == "shaanxi"))
par(mfrow = c(2,2))
plot(lm_mod_log2_cohort)


summary(lm_mod_log2_cohort)
broom::tidy(lm_mod_log2_cohort)
broom::glance(lm_mod_log2_cohort)
formula(lm_mod_log2_cohort)
broom::augment(lm_mod_log2_cohort)

```

```{r lm-lapply}
# run lms for each site individually, store in lists.

# log-log model, with cohort fixed effects
lm_log2_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0 + log(time_1):cohort,  
               data = filter(dat_l, site == site_names[i]))
})

# log-log model, without cohort fixed effects
lm_log2_no_cohort_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0 + log(time_1),  
               data = filter(dat_l, site == site_names[i]))
})

# log-log model, with year_abn (continuous) rather than the categorical cohort
lm_log2_year_abn_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0  + log(time_1) + log(time_1):year_abn,  
               data = filter(dat_l, site == site_names[i]))
})

# lon-linear (exp. decay)
lm_log_lin_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0 + time:cohort,  
               data = filter(dat_l, site == site_names[i]))
})

# linear
lm_0_l <- lapply(1:11, FUN = function(i) {
  lm(log(proportion) ~ 0 + time,  
               data = filter(dat_l, site == site_names[i]))
})

names(lm_log2_l) <- site_names
names(lm_log2_no_cohort_l) <- site_names
names(lm_log2_year_abn_l) <- site_names
names(lm_log_lin_l) <- site_names
names(lm_0_l) <- site_names
```


```{r plotting-one-site}
# ----------------------- log-log cohort plot ----------------------- #

# plot with age on the x-axis
ggplot(mapping = aes(x = age, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point(data = filter(dat_l, site == "shaanxi")) + 
  scale_color_distiller(palette = "Greens") + 
  scale_x_continuous(n.breaks = 30) + 
  labs(title = "Decay of abandoned land over time", #subtitle = site_df$description[indx], 
       caption = "lm formula: log(proportion) ~ 0 + log(time_1):cohort",
       x = "Time since initial abandonment (years)", y = "Proportion of abandoned land remaining",
       color = "Cohort (Year \nFirst Abandoned)") + 
  
  # add the lm curves, with predictions (this works because the "newdata" df has "age" in it, as well as "time_1")
  geom_line(data = mutate(augment(lm_mod_log2_cohort, newdata = filter(dat_l_0, site == "shaanxi")),
                          proportion = exp(.fitted))) + 
  
  # add the mean coefficient (needs to be transformed 4 to the right, so that plotting works)
  geom_function(fun = function(x) { (x-4) ^ (mean(coef(lm_log2_l$shaanxi), na.rm = TRUE))},
                mapping = aes(linetype = "Mean Decay Rate"),
                color = "blue", size = 1, #linetype = "dashed", 
                inherit.aes = FALSE) +
  scale_linetype(name = "")



# ----------------------------- #
# plot with time_1 on the x-axis

ggplot(data = filter(dat_l, site == "shaanxi"), 
       mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_line(data = mutate(augment(lm_mod_log2_cohort), # or augment(lm_log2_l$shaanxi)
                          time_1 = exp(`log(time_1)`),
                          proportion = exp(.fitted),
                          year_abn = as.numeric(levels(cohort))[cohort])) + 
  scale_x_continuous(n.breaks = 30) + 
  
  # the mean coefficient
  geom_function(fun = function(x) { x ^ mean(coef(lm_mod_log2_cohort), na.rm = TRUE)},
                size = 1, linetype = "dashed", inherit.aes = FALSE)

```



```{r plotting-diagnostics}

# ----------------------- diagnostics plot, for all 11 sites in sequence ----------------------- #
lapply(1:11, FUN = function(i) {
  
  png(filename = paste0(your_path, "log2_cohort_diag_", names(lm_log2_l[i]),".png"), 
    width = 8, height = 8, units = "in", res = 400)
  par(mfrow = c(2,2))
  plot(lm_log2_l[[i]], 1, main = site_names[i])
  plot(lm_log2_l[[i]], 2)
  plot(lm_log2_l[[i]], 3)
  plot(lm_log2_l[[i]], 4)

  dev.off()
  
})


# ----------------------- fitted vs. residuals diag plot, all 11 sites in one figure ----------------------- #
png(filename = paste0(your_path, "log2_cohort_diag_resid_v_fitted.png"), 
    width = 9, height = 8, units = "in", res = 400)
par(mfrow = c(3, 4))
for (i in 1:11) plot(lm_log2_l[[i]], 1, main = site_names[i])
dev.off()

```


```{r single-lm-all-sites}
# construct a model of *all* sites, with cohort and site fixed effects
str(dat_l)

# log-log model, with cohort and site fixed effects
lm_log2_cohort_all_site <- lm(log(proportion) ~ 0 + log(time_1):cohort:site, data = dat_l)


# ------------------------------------------------ #
# diagnostic plots
# ------------------------------------------------ #
par(mfrow = c(2,2))
plot(lm_log2_cohort_all_site, main = "Model with all sites, all cohorts")

# now, without the outlier
par(mfrow = c(2,2))
plot(lm(log(proportion) ~ 0 + log(time_1):cohort:site, data = dat_l[-2092, ]), main = "Model with all sites, all cohorts")

```


```{r log-time-linear-time}
# now with a log(time) term and a linear time term: 
lm_mod_log2_cohort_plus <- lm(log(proportion) ~ 0 + log(time_1):cohort + time_1:cohort, data = filter(dat_l, site == "shaanxi"))

ggplot(data = filter(dat_l, site == "shaanxi"), 
       mapping = aes(x = time_1, y = proportion, group = year_abn, color = year_abn)) + 
  geom_point() + scale_color_distiller(palette = "Greens") + 
  geom_line(data = mutate(augment(lm_mod_log2_cohort_plus), # or augment(lm_log2_l$shaanxi)
                          time_1 = exp(`log(time_1)`),
                          proportion = exp(.fitted),
                          year_abn = as.numeric(levels(cohort))[cohort])) + 
  scale_x_continuous(n.breaks = 30)


par(mfrow = c(2, 2))
plot(lm_mod_log2_cohort_plus)



```

```{r resid-v-covariates}
# residuals vs. covariates
lm_mod_log2_cohort

par(mfrow = c(2,2))
plot(lm_mod_log2_cohort)
augment(lm_mod_log2_cohort) %>% arrange(desc(.hat))

# just one site
resid_df1 <-  bind_cols(augment(lm_mod_log2_cohort), # with the tester model from above
                        filter(dat_l, site == "shaanxi"))

# all sites individually, combined
resid_df <- lapply(lm_log2_l, augment) %>% bind_rows(.id = "site") %>%
  bind_cols(., dat_l)


# residuals from the full model including all sites
resid_df_all <- bind_cols(dat_l, 
                          augment(lm_log2_cohort_all_site))




# plot for just shaanxi
str(resid_df)
gg_resid_time <- ggplot(data = resid_df, 
                        aes(x = time, y = .std.resid, color = year_abn)) + 
  geom_point() + 
  scale_color_distiller(palette = "Greens") + 
  labs(y = "Residuals", x = "Time Past Abandonment Threshold")

gg_resid_area <- ggplot(data = resid_df, 
                        aes(x = initial_area_abn, y = .std.resid, color = year_abn)) + 
  geom_point(position = position_jitter(width = 20)) + 
  scale_color_distiller(palette = "Greens") + 
  labs(y = "Residuals", x = "Initial Area Abandoned") + 
  theme(legend.position = "none")



#install.packages("cowplot")
library(cowplot)
# just shaanxi
cowplot::plot_grid(gg_resid_time, gg_resid_area, rel_widths = c(1.2, 1))


# all sites
cowplot::plot_grid(
  gg_resid_time %+% resid_df[-2092, ], # without outlier
  gg_resid_area %+% resid_df[-2092, ], # without outlier
  rel_widths = c(1.2, 1))


# residuals from the model with all sites: 
cowplot::plot_grid(
  gg_resid_time %+% resid_df_all[-2092, ], # without outlier
  gg_resid_area %+% resid_df_all[-2092, ], # without outlier
  rel_widths = c(1.2, 1))



# to plot specific sites:
cowplot::plot_grid(
  gg_resid_time %+% filter(resid_df[-2092, ], site...1 == "shaanxi"), # without outlier
  gg_resid_area %+% filter(resid_df[-2092, ], site...1 == "shaanxi"), # without outlier
  rel_widths = c(1.2, 1))
```
